### 2011-09-07 pkg/partykit/inst/doc/partykit.Rnw
further improvements to vignette
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@220 255f4249-2263-4417-b1ea-8498779e27a7

\documentclass[nojss]{jss}

%\VignetteIndexEntry{partykit: A Toolkit for Recursive Partytioning}
%\VignetteDepends{partykit}
%\VignetteKeywords{recursive partitioning, regression trees, classification trees, decision trees}
%\VignettePackage{partykit}

%% packages
\usepackage{amstext}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{thumbpdf}
\usepackage{rotating}
%% need no \usepackage{Sweave}

%% additional commands
\newcommand{\squote}[1]{`{#1}'}
\newcommand{\dquote}[1]{``{#1}''}
\newcommand{\fct}[1]{{\texttt{#1()}\index{#1@\texttt{#1()}}}}
\newcommand{\class}[1]{\dquote{\texttt{#1}}}

%% further commands
\renewcommand{\Prob}{\mathbb{P} }
\renewcommand{\E}{\mathbb{E}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\Var}{\mathbb{V}}
\newcommand{\R}{\mathbb{R} }
\newcommand{\N}{\mathbb{N} }
\newcommand{\C}{\mathbb{C} }
\newcommand{\argmin}{\operatorname{argmin}\displaylimits}
\newcommand{\argmax}{\operatorname{argmax}\displaylimits}
\newcommand{\LS}{\mathcal{L}_n}
\newcommand{\TS}{\mathcal{T}_n}
\newcommand{\LSc}{\mathcal{L}_{\text{comb},n}}
\newcommand{\LSbc}{\mathcal{L}^*_{\text{comb},n}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\A}{\mathcal{A}}
\newcommand{\yn}{y_{\text{new}}}
\newcommand{\z}{\mathbf{z}}
\newcommand{\X}{\mathbf{X}}
\newcommand{\Y}{\mathbf{Y}}
\newcommand{\sX}{\mathcal{X}}
\newcommand{\sY}{\mathcal{Y}}
\newcommand{\T}{\mathbf{T}}
\newcommand{\x}{\mathbf{x}}
\renewcommand{\a}{\mathbf{a}}
\newcommand{\xn}{\mathbf{x}_{\text{new}}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\ws}{\mathbf{w}_\cdot}
\renewcommand{\t}{\mathbf{t}}
\newcommand{\M}{\mathbf{M}}
\renewcommand{\vec}{\text{vec}}
\newcommand{\B}{\mathbf{B}}
\newcommand{\K}{\mathbf{K}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\I}{\mathbf{I}}
\newcommand{\bS}{\mathbf{S}}
\newcommand{\cellx}{\pi_n[\x]}
\newcommand{\partn}{\pi_n(\mathcal{L}_n)}
\newcommand{\err}{\text{Err}}
\newcommand{\ea}{\widehat{\text{Err}}^{(a)}}
\newcommand{\ecv}{\widehat{\text{Err}}^{(cv1)}}
\newcommand{\ecvten}{\widehat{\text{Err}}^{(cv10)}}
\newcommand{\eone}{\widehat{\text{Err}}^{(1)}}
\newcommand{\eplus}{\widehat{\text{Err}}^{(.632+)}}
\newcommand{\eoob}{\widehat{\text{Err}}^{(oob)}}
\newcommand{\bft}{\mathbf{t}}

\hyphenation{Qua-dra-tic}

\title{\pkg{partykit}: A Toolkit for Recursive Partytioning}
\Plaintitle{partykit: A Toolkit for Recursive Partytioning}

\author{Torsten Hothorn\\Ludwig-Maximilians-Universit\"at \\ M\"unchen
   \And Achim Zeileis\\Universit\"at Innsbruck}
\Plainauthor{Torsten Hothorn, Achim Zeileis}

\Abstract{
  This vignette is intended to be a short technical introduction to the \pkg{partykit} package.
  It is still unfinished but hopefully already helpful to some interested users.

  The purpose of the package is to provide a toolkit with infrastructure for representing,
  summarizing, and visualizing tree-structured regression and classification models.
  Thus, the focus is not on \emph{inferring} such a tree structure from data but
  to \emph{represent} a given tree so that print/plotting and computing predictions
  can be performed in a standardized way. In particular, this unified infrastructure can be
  used for reading/coercing tree models from different sources
  (\pkg{rpart}, \pkg{RWeka}, PMML) yielding objects that share functionality for
  \fct{print}/\fct{plot}/\fct{predict} methods.
}
\Keywords{recursive partitioning, regression trees, classification trees, decision trees}

\Address{
  Torsten Hothorn\\
  Institut f\"ur Statistik\\
  Ludwig-Maximilians-Universit\"at M\"unchen\\
  Ludwigstr.~33\\
  80539 M\"unchen, Germany\\
  E-mail: \email{Torsten.Hothorn@R-project.org}\\
  URL: \url{http://www.stat.uni-muenchen.de/~hothorn/}\\

  Achim Zeileis\\
  Department of Statistics \\
  Universit\"at Innsbruck \\
  Universit\"atsstr.~15 \\
  6020 Innsbruck, Austria \\
  E-mail: \email{Achim.Zeileis@R-project.org}\\
  URL: \url{http://eeecon.uibk.ac.at/~zeileis/}
}

\begin{document}

\SweaveOpts{eps=FALSE, keep.source=TRUE, eval = TRUE}

<<setup, echo = FALSE, results = hide>>=
options(width = 70)
library("partykit")
set.seed(290875)
data("iris")
@

\section{Motivating example and overview} \label{sec:intro}

To illustrate how \pkg{partykit} can be used to represent trees,
a simple artificial data set from Witten \& Frank's book
\emph{Data Mining: Practical Machine Learning Tools and Techniques}
is used. It concerns the conditions suitable for playing some
unspecified game.
<<weather-data>>=
data("WeatherPlay", package = "partykit")
WeatherPlay
@
To represent the \code{play} decision based on the corresponding
weather condition variables one could use the tree displayed in
Figure~\ref{weather-plot}. For now, it is ignored how this tree
was \emph{inferred} and it is simply assumed to be given.

\setkeys{Gin}{width=0.8\textwidth}
\begin{figure}[t!]
\centering
<<weather-plot0, echo=FALSE, fig=TRUE, height=5, width=7.5>>=
py <- party(
  partynode(1L,
    split = partysplit(1L, index = 1:3),
    kids = list(
      partynode(2L,
        split = partysplit(3L, breaks = 75),
        kids = list(
          partynode(3L, info = "yes"),
          partynode(4L, info = "no"))),
      partynode(5L, info = "yes"),
      partynode(6L,
        split = partysplit(4L, index = 1:2),
        kids = list(
          partynode(7L, info = "yes"),
          partynode(8L, info = "no"))))),
  WeatherPlay)
plot(py)
@
\caption{\label{weather-plot} Decision tree for \code{play} decision based on
  weather conditions in \code{WeatherPlay} data.}
\end{figure}
\setkeys{Gin}{width=\textwidth}

To represent this tree (also known as recursive partition) in \pkg{partykit},
two basic building blocks are used: splits of class \class{partysplit} and
nodes of class \class{partynode}. The resulting recursive partition can then
be associated with a data set in an object of class \class{party}.

\subsubsection*{Splits}

First, we employ the \fct{partysplit} function to creat the three
splits in the ``weather tree'' from Figure~\ref{weather-plot}. The
function takes the following arguments
\begin{Code}
  partysplit(varid, breaks = NULL, index = NULL, ..., info = NULL)
\end{Code}
where \code{varid} is an integer id of the variable used for splitting
e.g., \code{1L} for \code{outlook}, \code{3L} for \code{humidity},
\code{4L} for \code{windy} etc. Then, \code{breaks} and \code{index}
determine which observations are sent to which of the branches, e.g.,
\code{breaks = 75} for the humidity split. Apart from further arguments
not shown above (and just comprised under `\code{...}'), some arbitrary
information can be associated with a \class{partysplit} object by passing
it to the \code{info} argument. The three splits from Figure~\ref{weather-plot}
can then be created via
<<weather-splits>>=
sp_o <- partysplit(1L, index = 1:3)
sp_h <- partysplit(3L, breaks = 75)
sp_w <- partysplit(4L, index = 1:2)
@
For the numeric \code{humidity} variable the \code{breaks} are set while
for the factor variables \code{outlook} and \code{windy} the information
is supplied which of the levels should be associated with which of the
branches of the tree.

\subsubsection*{Nodes}

Second, we use these splits in the creation of the whole decision tree.
In \pkg{partykit} a tree is represented by a \class{partynode} object
which is recursive in that it may have ``kids'' that are again \class{partynode}
objects. These can be created with the function
\begin{Code}
  partynode(id, split = NULL, kids = NULL, ..., info = NULL)
\end{Code}
where \code{id} is an integer identifier of the node number, \code{split}
is a \class{partysplit} object, and \code{kids} is a list of \class{partynode}
objects. Again, there are further arguments not shown (\code{...}) and
arbitrary information can be supplied in \code{info}. The whole tree
from Figure~\ref{weather-plot} can then be created via
<<weather-nodes>>=
pn <- partynode(1L, split = sp_o, kids = list(
  partynode(2L, split = sp_h, kids = list(
    partynode(3L, info = "yes"),
    partynode(4L, info = "no"))),
  partynode(5L, info = "yes"),
  partynode(6L, split = sp_w, kids = list(
    partynode(7L, info = "yes"),
    partynode(8L, info = "no")))))
@
where the previously created \class{partysplit} objects are used as splits
and the nodes are simply numbered from~1 to~8. For the terminal nodes of the
tree there are no \code{kids} and the corresponding \code{play} decision
is stored in the \code{info} argument. Printing the \class{partynode} object
reflects the recursive structure stored.
<<weather-nodes-print>>=
pn
@
However, the displayed information is still rather raw as it has not yet
been associated with the \code{WeatherPlay} data set.

\subsubsection*{Trees (or recursive partitions)}

Hence, in a third step the recursive tree structure stored
in \code{pn} is coupled with the corresponding data in a \class{party}
object.
<<weather-party>>=
py <- party(pn, WeatherPlay)
py
@
And Figure~\ref{weather-plot} can easily be created by
<<weahter-plot, eval=FALSE>>=
plot(py)
@
In addition to the \class{partynode} and the \class{data.frame},
the function \fct{party} takes several further arguments 
\begin{Code}
  party(node, data, fitted = NULL, terms = NULL, ..., info = NULL)
\end{Code}
i.e., \code{fitted} values, a \code{terms} object, arbitrary additional
\code{info}, and again some further arguments comprised in \code{...}.

\subsubsection*{Further information}

More detailed technical information is provided in the subsequent sections.
Section~\ref{sec:splits}, \ref{sec:nodes}, and~\ref{sec:trees} discuss
the \fct{partysplit}, \fct{partynode}, and \fct{party} functions, respectively.
Section~\ref{sec:application} illustrates how this infrastructure can
be employed in a function that recursively \emph{infers} a tree.

\subsubsection{Design principles}

To facilitate reading of the subsequent sections, two design principles
employed in the creation of \pkg{partykit} are briefly explained.

(1)~Many helper utilities are encapsulated in functions that follow
a simple naming convention. To extract/compute some information \emph{foo}
from splits, nodes, or trees, \pkg{partykit} provides
\emph{foo}\code{_split}, \emph{foo}\code{_node}, \emph{foo}\code{_party}
functions (that are applicable to \class{partysplit}, \class{partynode},
and \class{party} objects, repectively). An example for the information
\emph{foo} might be \code{kidids}. Such functions are typically
not to be called by the end-user but potentially by package designers that
want to build functionality on top of \pkg{partykit}.

(2)~Printing and plotting relies on \emph{panel functions} that visualize
and/or format certain aspects of the resulting display, e.g., that of inner nodes,
terminal nodes, headers, footers, etc. A simple example would be printing
with a custom panel function for formatting the terminal node:
<<weather-party-print>>=
print(py,
  terminal_panel = function(node) paste(": play=", node$info, sep = ""))
@
Furthermore, arguments like \code{terminal_panel} can also take
\emph{panel-generating functions}, i.e., functions that produce a panel
function when applied to the \class{party} object.

\section{Splits} \label{sec:splits}

A split is basically a function that maps data, 
more specifically a partitioning variable, to daugther nodes. 
Objects of class \class{partysplit} are designed to represent 
such functions and are set-up by the \fct{partysplit} 
constructor:
<<partysplit-1, echo = TRUE>>=
## binary split in numeric variable `Sepal.Length'
sl5 <- partysplit(which(names(iris) == "Sepal.Length"), breaks = 5)
class(sl5)
@
The internal structure of class \class{partysplit} contains
information about the partitioning variable, the split-points,
the handling of split-points, the treatment of observations with
missing values and the daughter nodes to send observations to:
<<partysplit-2, echo = TRUE>>=
unclass(sl5)
@
Here, the split is defined in the first variable (corresponds to
\code{Sepal.Length} in data frame \code{iris}) and the splitting rule
is \code{Sepal.Length} $\le 5$:
<<partysplit-3, echo = TRUE>>=
character_split(sl5, data = iris)
@
This representation of splits is completely abstract and, most importantly,
independent of any data. Now, data comes into play when we actually
want to perform splits:
<<partysplit-4, echo = TRUE>>= 
kidids_split(sl5, data = iris)
@
For each observation in \code{iris} the split is performed and the 
number of the daughter node to send this observation to
is returned. Of course, this is a very complicated way of saying
<<partysplit-5, echo = TRUE>>=
(!with(iris, Sepal.Length <= 5)) + 1
@

Formally, a split is a function $f$ mapping an element $x = (x_1, \dots, x_p)$ of a $p$-dimensional 
sample space $\mathcal{X}$ into a set of $k$ daugther nodes $\mathcal{D} = \{d_1, \dots, d_k\}$.
This mapping is defined as a composition $f = h \circ g$ of two functions $g: \mathcal{X} \rightarrow \mathcal{I}$
and $h: \mathcal{I} \rightarrow \mathcal{D}$ with index set $\mathcal{I} = \{1, \dots, l\}, l \ge k$.

Let $\mu = (-\infty, \mu_1, \dots, \mu_{l - 1}, \infty)$ denote the split points
($(\mu_1, \dots, \mu_{l - 1})$ = \code{breaks}). We are interested to split according to the 
information contained in the $i$th element of $x$ ($i$ = \code{varid}).
For numeric $x_i$, the split points are also numeric. If $x_i$ is a factor at
levels $1, \dots, K$, the default split points are $\mu = (-\infty, 1, \dots, K - 1, \infty)$.

The function $g$ essentially determines, which of the intervals (defined by $\mu$) the value
$x_i$ is contained in ($I$ denotes the indicator function here):
\begin{eqnarray*}
x \mapsto g(x) = \sum_{j = 1}^l j I_{\mathcal{A}(j)}(x_i)
\end{eqnarray*}
where $\mathcal{A}(j) = (\mu_{j - 1}, \mu_j]$ for \code{right = TRUE} except 
$\mathcal{A}(l) = (\mu_{l - 1}, \infty)$.
If \code{right = FALSE}, then $\mathcal{A}(j) = [\mu_{j - 1}, \mu_j)$ except 
$\mathcal{A}(1) = (-\infty, \mu_1)$. Note that with some categorical variable $x_i$ and default
split points, $g$ is the identity.

Now, $h$ maps from the index set $\mathcal{I}$ into the set of daugther nodes:
\begin{eqnarray*}
f(x) = h(g(x)) = d_{\sigma_{g(x)}}
\end{eqnarray*}
where $\sigma = (\sigma_1, \dots, \sigma_l) \in \{1, \dots, k\}^l$ (\code{index}). 
By default, $\sigma = (1, \dots, l)$ and $k = l$.

If $x_i$ is missing, then $f(x)$ is randomly drawn with $\mathbb{P}(f(x) = d_j) = p_j, j = 1, \dots, k$
for a discrete probability distribution $p = (p_1, \dots, p_k)$ over the $k$ daugther nodes (\code{prob}).

In the simplest case of a binary split in a numeric variable $x_i$, there is only
one split point $\mu_1$ and, with $\sigma = (1, 2)$, observations with $x_i \le \mu_1$ are sent
to daugther node $d_1$ and observations with $x_i > \mu_1$ to $d_2$. However,
this representation of splits is general enough to deal with more complicated
set-ups like surrogate splits, where typically the index needs modification, for example
$\sigma = (2, 1)$, categorical splits, i.e., there is one data structure for both
ordered and unordered splits, multiway splits, and functional splits. The latter
can be implemented by defining a new artificial splitting variable $x_{p + 1}$ by
means of a potentially very complex function of $x$ later used for splitting.

As an example, consider a split in a categorical
variable at three levels where the first two levels go to the left
daugther node and the third one to the right daugther node:
<<partysplit-6, echo = TRUE>>=
## binary split in factor `Species'
sp <- partysplit(which(names(iris) == "Species"), index = c(1L, 1L, 2L))
character_split(sp, data = iris)
table(kidids_split(sp, data = iris), iris$Species)
@
The internal structure of this object contains the \code{index} slot
<<partysplit-6, echo = TRUE>>=
unclass(sp)
@
that maps levels to daughter nodes. This mapping is also
useful with splits in ordered variables, for example when 
representing multiway splits:
<<partysplit-7, echo = TRUE>>=
## multiway split in numeric variable `Sepal.Width',    
## higher values go to the first kid, smallest values
## to the last kid
sw23 <- partysplit(which(names(iris) == "Sepal.Width"),
  breaks = c(3, 3.5), index = 3:1)
character_split(sw23, data = iris)
table(kidids_split(sw23, data = iris),
  cut(iris$Sepal.Width, breaks = c(-Inf, 2, 3, Inf)))
@
The mapping of classes of the categorized numeric variable
to daugther nodes can be changed by modifying \code{index}:
<<partysplit-8, echo = TRUE>>=
sw23 <- partysplit(which(names(iris) == "Sepal.Width"),
  breaks = c(3, 3.5), index = c(1L, 3L, 2L))
character_split(sw23, data = iris)
@

The additional argument \code{prop} is used to specify a 
discrete probability distribution over the daugther nodes that
is used to map observations with missing values to daugther nodes.
Furthermore, the \code{info} argument and slot takes arbitrary objects
to be stored with the split (for example split statistics) but is not
structured at the moment.

The slots of \class{partysplit} objects shall be accessed by the corresponding
accessor functions.

\section{Nodes} \label{sec:nodes}

Inner and terminal nodes are represented by objects of class \class{partynode}.
Each node has a unique identifier \code{id}. A node consisting only
of such an identifier (and possibly additional information  
in \code{info}) is a terminal node:
<<partynode-1, echo = TRUE>>=
n1 <- partynode(id = 1L)
is.terminal(n1)
print(n1)
@
Inner nodes have to have a primary split \code{split} and at least two
daugther nodes. The daugther nodes are objects of class \class{partynode}
itself and thus represent the recursive nature of this data structure.
The daugther nodes are pooled in a list \code{kids}. 
In addition, a list of \class{partysplit} objects offering 
surrogate splits can be supplied; a list of \class{partysplit} objects
in slot \code{surrogates} defines such additional splits (mostly used 
for handling missing values). Note that \class{partynode} objects aren't 
connected to the actual data.

Based on the binary split \code{sl5} defined in the previous section, 
we set-up an inner node with two terminal daugther nodes and
print this stump (the data is needed because neither split nor nodes
contain information about variable names or levels):
<<partynode-2, echo = TRUE>>=
n1 <- partynode(id = 1L, split = sl5, kids = sapply(2:3, partynode))
print(n1, data = iris)
@
Now that we have defined our first simple tree, we want to assign
observations to terminal nodes:
<<partynode-3, echo = TRUE>>=
fitted_node(n1, data = iris)
@
Here, the \code{id}s of the terminal node each observations falls into
are returned. Alternatively, we could compute the position of these
daugther nodes in the list \code{kids}:
<<partynode-4, echo = TRUE>>=
kidids_node(n1, data = iris)
@
Furthermore, the \code{info} argument and slot takes arbitrary objects
to be stored with the node (predictions, for example, but we will handle
this issue later). The slots can be extracted by means of the corresponding
accessor functions.

A number of methods are defined for \class{partynode} objects.
\fct{is.partynode} checks if the argument is a valid \class{partynode}
object. \fct{is.terminal} is \code{TRUE} for terminal nodes
and \code{FALSE} for inner nodes. The subset methods
return the \class{partynode} object corresponding to the \code{i}th
kid:
<<partynode-5, echo = TRUE>>=
n1[2]
@

The \fct{as.partynode} and \fct{as.list} methods can be used
to convert flat list structures into recursive \class{partynode}
objects and vice versa. \fct{as.partynode} applied to
\class{partynode} objects renumbers the recursive nodes
starting with root node identifier \code{from}.

\fct{length} gives the number of kid nodes of the root node,
\fct{depth} the depth of the tree and \fct{width}
the number of terminal nodes.

\section{Trees} \label{sec:trees}

%%% code in eine Tabelle

Although tree structures can be represented by \class{partynode} objects,
a tree is more than a number of nodes and splits. More
information about (parts of the) corresponding data is necessary for 
high-level computations on trees.

Objects of class \class{party} basically consist of a \class{partynode}
object representing the tree structure in a recursive way and 
data. The \code{data} argument takes a \class{data.frame} which, however,
might have zero columns. Optionally, a \class{data.frame} with at least one
variable \code{(fitted)} containing the terminal node numbers of
data used for fitting the tree may be specified along with a   
\code{terms} object or any additional (currently unstructured)
information as \code{info}. Argument \code{names} defines names
for all nodes in \code{node}.

<<party-1, echo = TRUE>>=
t1 <- party(n1, 
  data = iris,
  fitted = data.frame(
    "(fitted)" = fitted_node(n1, data = iris),
    "(response)" = iris$Species,
    check.names = FALSE)
)
t1
@

\section{My first tree} \label{sec:application}

Package \pkg{partykit} does not offer unified infrastructure for growing trees. However,
once you know how to estimate splits from data, it is fairly straightforward to
implement trees. Consider a very simple tree algorithm. We assume that both response
and features are numeric. We search for the binary best split by means of $t$-test 
$p$-values, i.e., we cycle through all variables and potential split points and assess
the quality of the split by comparing the distributions of the response in the so-defined
two groups. We select the feature/split point combination with lowest two-sided $p$-value,
however only if this result is significant at level $\alpha = 0.05$.

This strategy can be implemented based on the data (response and features) and 
some case weights as follows (\code{response} is just the name of the response
and \code{data} is a data frame with all variables):
<<mytree-1, echo = TRUE>>=
findsplit <- function(response, data, weights) {

  ### extract response values from data
  y <- data[[response]]

  logpmin <- 0
  xselect <- NULL

  ### cycle through all features
  for (i in which(names(data) != response)) {

    ### expand data
    x <- data[[i]]
    xt <- rep(x, weights)
    yt <- rep(y, weights)

    ### potential split points (not too many)
    qx <- unique(quantile(xt, 
        	 prob = seq(from = 0.1, to = 0.9, by = 0.05)))

    ### assess all potential splits by their t-test
    ### log-p-value
    logp <- sapply(qx, function(q) {
      tt <- t.test(yt[xt <= q], yt[xt > q])
      pt(-abs(tt$statistic), tt$parameter, log = TRUE) + log(2)
    })

    ### if the best split in variable i significant AND
    ### better than what we already had store this information
    if (min(logp) < logpmin & min(logp) < log(0.05)) {
      logpmin <- min(logp)
      xselect <- i
      splitpoint <- qx[which.min(logp)]
    }
  }

  ### no significant split found, give up
  if (is.null(xselect)) return(NULL)

  ### return split as partysplit object
  return(partysplit(
      varid = as.integer(xselect),	 ### which variable?
      breaks = as.numeric(splitpoint),   ### which split point?
      info = list(pvalue = exp(logpmin)  ### save p-value in addition
  )))
}
@

In order to actually grow a tree on data, 
we have to set-up the recursion for growing a recursive 
\class{partynode} structure:
<<mytree-2, echo = TRUE>>=
growtree <- function(id = 1L, response, data, weights) {

  ### for less than 30 obs. stop here
  if (sum(weights) < 30) return(partynode(id = id))

  ### find best split
  sp <- findsplit(response, data, weights)
  ### no split found, stop here
  if (is.null(sp)) return(partynode(id = id))

  ### actually split the data
  kidids <- kidids_split(sp, data = data)

  ### set-up all daugther nodes
  kids <- vector(mode = "list", length = max(kidids))
  for (kidid in 1:max(kidids)) {
  ### select obs for current node
  w <- weights
  w[kidids != kidid] <- 0
  ### get next node id
  if (kidid > 1) {
    myid <- max(nodeids(kids[[kidid - 1]]))
  } else {
    myid <- id
  }
  ### start recursion on this daugther node
  kids[[kidid]] <- growtree(id = as.integer(myid + 1), response, data, w)
  }

  ### return nodes
  return(partynode(id = as.integer(id), split = sp, kids = kids))
}
@

A very rough sketch of formula-based user-interface needs
to set-up the data and call \fct{growtree}:
<<mytree-3, echo = TRUE>>=
mytree <- function(formula, data, weights = NULL) {

  ### name of the response variable
  response <- all.vars(formula)[1]
  ### data without missing values, response comes last
  data <- data[complete.cases(data), c(all.vars(formula)[-1], response)]
  ### data is numeric
  stopifnot(all(sapply(data, is.numeric)))

  if (is.null(weights)) weights <- rep(1, nrow(data))
  ### weights are case weights, i.e., integers
  stopifnot(length(weights) == nrow(data) &
    max(abs(weights - floor(weights))) < .Machine$double.eps)

  ### grow tree
  nodes <- growtree(id = 1L, response, data, weights)

  ### compute terminal node number for each obs.
  fitted <- fitted_node(nodes, data = data)
  ### return rich object
  ret <- party(nodes, 
    data = data,
    fitted = data.frame(
      "(fitted)" = fitted,
      "(response)" = data[[response]],
      "(weights)" = weights,
      check.names = FALSE),
    terms = terms(formula))
  as.constparty(ret)
}
@

We now can fit this tree, for example to the airquality data; the
\fct{print} method provides us with a first overview on the 
resulting model
<<mytree-4, echo = TRUE>>=
aqt <- mytree(Ozone ~ Solar.R + Wind + Temp, data = airquality)
aqt
@

\begin{figure}[t!]
\centering
<<mytree-5, echo = TRUE, fig = TRUE, width = 10, height = 6>>=
plot(aqt)
@
\caption{Tree. \label{plottree}}
\end{figure}

We depict the model graphically using \fct{plot} (see Figure~\ref{plottree})
and compute predictions using
<<mytree-6, echo = TRUE>>=
predict(aqt, newdata = airquality[1:10,])
@

An interesting feature is the ability to extract subsets of trees:
<<mytree-7, echo = TRUE>>=
aqt4 <- aqt[4]
aqt4
@
which again are objects inheriting from \class{party} and thus
can be plotted easily (see Figure~\ref{subtree}).

\begin{figure}[t!]
\centering
<<mytree-8, echo = TRUE, fig = TRUE, width = 10, height = 6>>=
plot(aqt4)
@
\caption{Subtree. \label{subtree}}
\end{figure}

We also might be interested in extracting the $p$-values in the
inner nodes in a nicely formatted way:
<<mytree-10, echo = TRUE>>=
fun <- function(x) format.pval(info_split(split_node(x))$pvalue,
  digits = 3, eps = 0.001)
nid <- nodeids(aqt)
iid <- nid[!(nid %in% nodeids(aqt, terminal = TRUE))]
unlist(nodeapply(aqt, ids = iid, FUN = fun))
@

\end{document}