### 2011-09-26 pkg/evtree/inst/doc/evtree.Rnw
Improved almost all sections of the paper, a few FIXMEs remain.
Compilation requires deleting the cached .rda first.



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@249 255f4249-2263-4417-b1ea-8498779e27a7

\documentclass[nojss]{jss}
\usepackage{amssymb, amsmath, amsthm, booktabs, thumbpdf}
\newcommand{\argmin}{\operatorname{argmin}\displaylimits}

%% for internal use
\newcommand{\fixme}[1]{\emph{\marginpar{FIXME} (#1)}}
\newcommand{\readme}[1]{\emph{\marginpar{README} (#1)}}

%% Sweave/vignette information and metadata
%% need no \usepackage{Sweave}
\SweaveOpts{engine = R, eps = FALSE, keep.source = TRUE}
%\VignetteIndexEntry{Evolutionary Learning of Globally Optimal Classification and Regression Trees in R}
%\VignetteDepends{stats,evtree,partykit,mlbench,multcomp,rpart,party,xtable}
%\VignetteKeywords{machine learning, classification trees, regression trees, evolutionary algorithms, R}
%\VignettePackage{evtree}

\author{Thomas Grubinger\\ Innsbruck Medical University 
   \And Achim Zeileis\\ Universit\"at Innsbruck 
   \And Karl-Peter Pfeiffer\\ Innsbruck Medical University
}
\Plainauthor{Thomas Grubinger, Achim Zeileis, Karl-Peter Pfeiffer}

\title{\pkg{evtree}: Evolutionary Learning of Globally Optimal Classification and Regression Trees in \proglang{R}}
\Plaintitle{evtree: Evolutionary Learning of Globally Optimal Classification and Regression Trees in R}
\Shorttitle{\pkg{evtree}: Evolutionary Learning of Globally Optimal Trees in \proglang{R}}

\Abstract{
  Commonly used classification and regression tree methods like the CART
  algorithm are recursive partitioning methods that build the model in a forward
  stepwise search. Although this approach is known to be an efficient heuristic,
  the results of recursive tree methods are only locally optimal, as splits are
  chosen to maximize homogeneity at the next step only. An alternative way to
  search over the parameter space of trees is to use global optimization methods
  like evolutionary algorithms. This paper describes the \pkg{evtree} package,
  which implements an evolutionary algorithm for learning globally optimal
  classification and regression trees in \proglang{R}. Computationally intensive
  tasks are fully computed in \proglang{C++} while the \pkg{partykit}
  \citep{hothorn2009partykit} package is leveraged for representing the resulting
  trees in \proglang{R}, providing unified infrastructure for summaries,
  visualizations, and predictions. \code{evtree} is compared to \code{rpart}
  \citep{therneau1997introduction}, the open-source CART implementation, and
  conditional inference trees \citep[\code{ctree},][]{hothorn2006urp}. The usefulness
  of \code{evtree} is illustrated in a textbook customer classification task
  and a benchmark study of predictive accuracy in which \code{evtree}
  achieved at least similar and most of the time better results compared to
  the recursive algorithms \code{rpart} and \code{ctree}.  
}

\Keywords{machine learning, classification trees, regression trees, evolutionary algorithms, \proglang{R}}
\Plainkeywords{machine learning, classification trees, regression trees, evolutionary algorithms, R}

\Address{
  Thomas Grubinger, Karl-Peter Pfeiffer\\
  Innsbruck Medical University\\
  Department for Medical Statistics, Informatics and Health Economics\\
  6020, Innsbruck, Austria\\
  E-mail: \email{Thomas.Grubinger@i-med.ac.at}, \email{Karl-Peter.Pfeiffer@i-med.ac.at}\\
  URL: \url{http://www.i-med.ac.at/msig/mitarbeiter/grubinger/}\\
  \phantom{URL: }\url{http://www.i-med.ac.at/msig/mitarbeiter/pfeiffer/}\\

  Achim Zeileis \\
  Universit\"at Innsbruck\\
  Department of Statistics\\
  6020, Innsbruck, Austria\\
  E-mail: \email{Achim.Zeileis@R-project.org}\\
  URL: \url{http://eeecon.uibk.ac.at/~zeileis/}
}

<<setup, echo=FALSE>>=
options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
library("rpart")
library("evtree")
data("BBBClub", package = "evtree")
cache <- TRUE
@

\begin{document}

\section{Introduction}

Classification and regression trees are commonly applied for exploration and
modeling of complex data. They are able to handle strongly nonlinear
relationships with high order interactions and different variable types. The
resulting model can be interpreted as a tree structure providing a compact
and intuitive representation. Commonly used classification and regression tree
algorithms, including \textit{CART} \citep{breiman1984cart} and \textit{C4.5}
\citep{quinlan1993}, use a greedy heuristic, where split rules are selected in a
forward stepwise search for recursively partitioning the data into groups. The
split rule at each internal node is selected to maximize the homogeneity of its
child nodes, without consideration of nodes further down the tree, hence yielding
only locally optimal trees. Nonetheless, the
greedy heuristic is computationally efficient and often yields reasonably good
results \citep{murthy1995dti}. However, for some problems, greedily induced
trees can be far from the optimal solution, and a global search over the tree's
parameter space can lead to much more compact and accurate models.

\pagebreak

The main challenge in growing globally optimal trees is that the search
space is typically huge rendering full-grid searches computationally infeasible.
One possibility to solve this problem is to use stochastic
optimization methods like evolutionary algorithms. 
In practice, however, such stochastic methods
are rarely used in decision tree induction. One reason is probably that they are
computationally much more demanding than a recursive forward search but another one is
likely to be the lack of availability in major software packages.
In particular, while there are several packages for \proglang{R} \citep{team2011r} providing
forward-search tree algorithms, there is only little support for globally
optimal trees. The former group of packages includes (among others)
\pkg{rpart} \citep{therneau1997introduction}, the open-source
implementation of the CART algorithm; \pkg{party}, containing two tree algorithms
with unbiased variable selection and statistical stopping criteria
\citep{hothorn2006urp,zeileis2008mob}; and \pkg{RWeka} \citep{rweka}, the
\proglang{R} interface to \pkg{Weka} \citep{weka} with open-source implementations
of tree algorithms such as C4.5 or \textit{M5} \citep{quinlan1992lcc}.
A notable exception is the
\pkg{LogicReg} package \citep{logicreg} for logic regression, an algorithm for globally optimal
trees based on binary covariates only  and using simulated annealing. See
\cite{hothorn2011ctv} for an overview of further recursive partitioning packages
for \proglang{R}.

To fill this gap, we introduce a new \proglang{R} package \pkg{evtree}, 
available from the Comprehensive \proglang{R} Archive Network at
\url{http://CRAN.R-project.org/package=evtree}, providing evolutionary
methods for learning globally optimal classification and regression trees.
Generally speaking, evolutionary algorithms are inspired by natural Darwinian
evolution employing concepts such as inheritance, mutation, and natural selection.
They are population-based, i.e., a whole collection of candidate solutions -- trees in this application -- is
processed simultaneously and iteratively modified by \textit{variation operators}
called \textit{mutation} (applied to single solutions) and \textit{crossover}
(merging different solutions). Finally, a survivor selection process favors
solutions that perform well according to some quality criterion, usually
called \textit{fitness function} or \textit{evaluation function}. In this
evolutionary process the mean quality of the population increases over time
\citep{baeck,eiben2003iec}.  
In the case of learning decision trees, this means that the variation
operators can be applied to modify the tree structure (e.g., number of splits, splitting
variables, and corresponding splitpoints etc.) in order to optimize a
fitness functions such as the misclassificaion or error rate penalized by
the complexity of the tree.
A notable difference to comparable algorithms is the survivor selection
mechanism where it is important to avoid premature convergence. In the
following, we use a simple $(1+1)$ selection strategy (i.e., one parent
solution competes with one offspring for a place in the population)
which can be argued to offer computational
advantages for the application to classification and regression trees.

The remainder of this paper is structured as follows: Section~\ref{got}
describes the problem of learning globally optimal decision trees and contrasts
it to the locally optimal forward-search heuristic that is utilized by recursive
partitioning algorithms. Section~\ref{desc} introduces the \code{evtree} algorithm
before Section~\ref{impl} addresses implementation details along with an
overview of the implemented functions. A benchmark comparison -- comprising
14~benchmark datasets, 3~real-world datasets, and 3~simulated scenarios -- is carried out
in Section~\ref{comp}, showing that the predictive performance of \code{evtree}
is often significantly better compared to the commonly used algorithms \code{rpart}
and \code{ctree} (from the \pkg{party} package). Finally, Section~\ref{conc} gives
concluding remarks about the implementation and the performance of the new
algorithm.

\section{Globally and locally optimal decision trees}
\label{got}

Classification and regression tree analysis aims at
modeling a response variable $Y$ by a vector of $P$ predictor variables
$X=(X_1,...,X_P)$ where for classification trees $Y$ is
qualitative and for regression trees $Y$ is quantitative. Tree-based methods
first partition the input space $X$ into a set of $M$ rectangular regions $R_m$
($m = 1,...,M$) then fit a (typically simple) model within each region $\{Y | X \in R_m\}$,
e.g., the mean, median, or variance etc. Typically, the mode is used for
classification trees and the arithmetic mean is applied for regression trees.

To show why forward-search recursive partitioning algorithms typically
lead to globablly suboptimal solutions, their parameter spaces and optimization
problems are presented and contrasted in a unified notation. Although all
arguments hold more generally, only binary tree models with some maximum number
of splits $M_{\max}$ are considered. Both restrictions make the notation somewhat
simpler while not really restricting the problem: (a)~Multiway splits are equivalent
to a sequence of binary splits in predictions and number of resulting subsamples.
(b)~The maximal size of the tree is always limited by the number of observations
in the learning sample.

In the following, a binary tree model with $M$ terminal nodes (which consequently has
$M-1$ internal splits) is denoted by
%
\begin{equation}
  \theta ~=~ (v_{n_1}, s_{n_1},...,v_{n_{M-1}},s_{n_{M-1}} ), \label{eq:tree}
\end{equation}
%
where the $n_r \in \{1,...,M_{\max}-1\}$ are the positions of the
internal \emph{nodes}, $v_r \in \{1, \dots ,P\}$ the associated splitting
\emph{variables}, and $s_r$ the associated \emph{split} rules ($r = 1, \dots, M-1$).
Depending on the domain of $X_{v_r}$, the split rule $s_r$ contains either a cutoff
(for ordered and numeric variables) or a a nonempty subset of $\{1, \dots, c\}$
(for a categorical variable with $c$~levels), determining which observations are
sent to the first or second subsample. In the former case, there are $u-1$ possible
split rules if $X_{v_r}$ takes $u$ distinct values; and in the latter case, there
are $2^{c-1} - 1$ possible splits. Thus, the product of all of these combinations
forms all potential elements $\theta$ from $\Theta_M$, the space of conceivable
trees with $M$ terminal nodes. The overall parameter space is then 
$\Theta =\bigcup_{M=1}^{M_{\max}} \Theta_{M}$ (which in practice is often reduced
by exluding elements $\theta$ resulting in too small subsamples etc.).

Finally, $f(X,\theta)$ denotes the prediction function based on all explanatory
variables $X$ and the chosen tree structure $\theta$ from Equation~\ref{eq:tree}.
As pointed out above, this is typically constructed using the means or modes in the respective
partitions of the learning sample.

\subsection{The parameter space of globally optimal decision trees}

As done by \cite{breiman1984cart}, let the complexity of a tree be measured by a
function of the number of terminal nodes, without further considering the depth
or the shape of trees. The goal is then to find that classification and regression tree
which optimizes some tradeoff between predicition performance and complexity:
%
\begin{equation} \label{eq:globally}
  \hat{\theta} ~=~ \argmin_{\theta \in \Theta} \; \mathrm{loss}\{Y, f(X,\theta)\} ~+~ \mathrm{comp}(\theta).
\end{equation}
%
where $\mathrm{loss}(\cdot, \cdot)$ is a suitable loss function for the
domain of $Y$; typically, the misclassification rate MC 
and the mean squared error MSE are employed for classification and regression,
respectively. The function $\mathrm{comp}(\cdot)$ is a function that is monotonically
non-decreasing in the number of terminal nodes $M$ of the tree $\theta$, thus penalizing
more complex models in the tree selection selection process.
Note that finding $\hat{\theta}$ requires a search over all $\Theta_M$.

The parameter space $\Theta$ becomes large for already medium sized problems and
a complete search for larger problems is computationally intractable. In fact,
\cite{hyafil1976cob} showed that building optimal binary decision trees, such
that the expected number of splits required to classify an unknown sample is
minimized, is NP-complete. \cite{zantema2000finding} proved that finding a
decision tree of minimal size that is decision-equivalent to a given decision
tree is also NP-hard. As a consequence the search space is usually limited
by heuristics.

\subsection{The parameter space of locally optimal decision trees}

Instead of searching all combinations in $\Theta$ simultaneously, recursive
partitioning algorithms only consider one split at a time. At each internal node
$r \in \{n_1,..., n_{M-1}\}$, the split variable $v_r$ and the corresponding
split point $s_r$ are selected to locally minimize the loss function.
Starting with an empty tree $\theta_0 = (\emptyset)$, the tree is first grown 
recursively for and subsequently \emph{pruned} to satisfy the complexity tradeoff:
%
\begin{eqnarray} \label{eq:locally}
  \bar \theta_r
    & = & \argmin_{ \theta = \theta_{r-1} \cup (v_r, s_r)} \; \mathrm{loss}\{Y, f(X, \theta)\} 
    \qquad (r = 1, \dots, M_{\max} -1), \\
  \bar \theta
    & = & \argmin_{\bar \theta_r} \; \mathrm{loss}\{Y, f(X, \bar \theta_r)\} ~+~ \mathrm{comp}(\bar \theta_r).
  \label{eq:pruning}
\end{eqnarray}
%
For nontrivial problems, forward-search recursive partitioning methods only search a small
fraction of the global search space $(v_1, s_1, \dots, v_{M_{\max}-1},
s_{M_{\max}-1} )$. They only search each $(v_r, s_r)$ once, and independently of
the subsequent split rules, hence typically leading to a globally suboptimal solution
$\bar \theta$.

Note that the notation above uses an exhaustive search for the $r$-th split,
jointly over $(v_r, s_r)$, as is employed in CART or C4.5. So-called \emph{unbiased}
recursive partitioning techniques modify this search by first seleting the variable $v_r$ using
statistical significance tests and subsequently selecting the optimal split $s_r$ for that
particular variable. This approach is used in conditional inference trees
\citep[see][for references to other algorithms]{hothorn2006urp} and avoids selecting
variables with many potential splits more often than those with fewer potential
splits.

\subsection{An illustration of the limitations of locally optimal decision trees}

A very simple example that illustrates the limitation of forward-search recursive partitioning
methods is depicted in Figure~\ref{fig:illustration}. The example only contains
two independent variables and can be solved with three splits that
partition the input space into four regions. As expected the recursive
partitioning methods \code{rpart} and \code{ctree} fail to find any split at
all, as the loss function on the resulting subsets cannot be reduced by the first
split. For methods that explore $\Theta$ in a more global fashion it is
straightforward to find an optimal solution to this problem. One solution
is the tree constructed by \code{evtree}: 
%
<<chess22, echo=FALSE>>=
X1 <- rep(seq(0.25, 1.75, 0.5), each = 4)
X2 <- rep(seq(0.25, 1.75, 0.5), 4)
Y <- rep(1, 16)
Y[(X1 < 1 & X2 < 1) | (X1 > 1 & X2 > 1)] <- 2
Y <- factor(Y, labels = c("O", "X"))
chess22 <- data.frame(Y, X1, X2)
print(evtree(Y ~ ., data = chess22, minbucket = 1, minsplit = 2))
@

\setkeys{Gin}{width=0.5\textwidth}
\begin{figure}[t!]
\centering
<<chess22-plot, fig=TRUE, echo=FALSE, height=4, width=4>>=
plot(X2 ~ X1, data = chess22, xlim = c(0, 2), ylim = c(0, 2), pch = c(1, 4)[Y], col = c("black", "slategray")[Y])
@
\caption{\label{fig:illustration} Class distribution of  the $(X_1, X_2)$-plane. The two classes are indicated by black circles and gray crosses.}
\end{figure}

All instances are classified correctly. Each of the terminal nodes~3 and~7 contain four instances of the class \code{X}. Four instances of class
\code{O} are assigned to each of the terminal nodes~4 and~6.

\subsection{Approaches for learning globally optimal decision trees}

When compared with the described forward stepwise search, a less greedy approach
is to calculate the effects of the split rules deeper down in the
tree. In this way optimal trees can be found for simple problems. However,
split selection at a given node in Equation~\ref{eq:locally} has complexity  $O(PN)$
(if all $P$ variables are numeric/ordered with $N$ distinct values).
Through a global search up to $M-1$ levels (\ref{eq:globally}) the complexity
increases to $O(P^M N^M)$ \citep{papagelis2001bdt}. One conceivable compromise
between these two extremes is to look ahead $d$~steps with $1 < d < M$
\citep[see e.g.,][]{esmeir2007ald}, also yielding a locally optimal tree but
less constrained than that from a 1-step-ahead search.
Another class of algorithms is given by stochastic optimization methods that,
given an initial tree, seek improved solutions through stochastic changes to
the tree structure. Thus, these algorithms try to explore the full parameter
space $\Theta$ but cannot be guaranteed to find the globally optimal solution
but only an approximation thereof. Besides evolutionary algorithms
\citep{koza1991concept}, \textit{Bayesian CART} \citep{denison1998bca} and
\textit{simulated annealing} \citep{sutton1992ict} were used successfully to
solve difficult classification and regression tree problems.
\cite{koza1991concept} first formulated the concept of using evolutionary
algorithms as a stochastic optimization method to build classification and
regression trees. \cite{papagelis2001bdt} presented a classification tree
algorithm and provided results on several datasets from the UCI machine learning
repository \citep{Frank+Asuncion:2010}. Another method for the construction of
classification and regression trees via evolutionary algorithms was introduced by
\cite{gray2008cta} and \cite{fan2005rta}, respectively.
\cite{cantu2003inducing} used an evolutionary algorithm to induce so-called oblique
classification trees.

\section[The evtree algorithm]{The \code{evtree} algorithm}
\label{desc}

The general framework of evolutionary algorithms emerged from different
representatives. \cite{Holland} called his method \textit{genetic algorithms},
\cite{rechenberg} invented \textit{evolution strategies}, and \cite{Fogel}
introduced \textit{evolutionary programming}. More recently, \cite{Koza}
introduced a fourth stream and called it \textit{genetic programming}. All four
representatives only differ in the technical details, for example the encoding
of the individual solutions, but follow the same general outline
\citep{eiben2003iec}. Evolutionary algorithms are being increasingly widely
applied to a variety of optimization and search problems. Common areas of
application include data mining \citep{freitas2002sea,cano2003uea}, statistics
\citep{de34glmulti}, signal and image processing \citep{man1997gac}, and
planning and scheduling \citep{jensen2001raf}.

\begin{table}[b!]
\centering
\hrulefill
\begin{enumerate}
  \item Initialize the population.
  \item Evaluate each individual.
  \item While(termination condition is not satisfied) do:
  \begin{itemize}
    \item[a.] Select parents.
    \item[b.] Alter selected individuals via variation operators.
    \item[c.] Evaluate new solutions.
    \item[d.] Select survivors for the next generation.
  \end{itemize}
\end{enumerate}
\hrulefill
\caption{Pseudocode of the general evolutionary algorithm.} 
\label{pc:genericEA}
\end{table}

The pseudocode for the general evolutionary algorithm is provided
in Table~\ref{pc:genericEA}. In the context of classification and regression trees,
all \emph{individuals} from the population (of some given size) are $\theta$s as defined in 
Equation~\ref{eq:tree}. The details of their evolutionary selection is
given below following the general outline displayed in Table~\ref{pc:genericEA}.

As pointed out in Section~\ref{got}, some elements $\theta \in \Theta$ are typically
excluded in practice to satisfy minimal subsample size requirements.
In the following, the term \emph{invalid node} refers to such excluded cases,
not meeting sample size restrictions.

\subsection{Initialization}
\label{init}

Each tree of the population is initialized with a valid, randomly generated,
split rule in the root node. First, $v_1$ is selected with uniform probability
from $1,...,P$. Second, a split point $s_1$ is selected. If $X_{v_1}$ is numeric
or ordinal with $u$ unique values, a split point $s_1$ is selected with uniform
probability from the $u-1$ possible spit points of $X_{v_1}$. If $X_{v_1}$ is
nominal and has $c$ categories, each $k=1,...,c$ has a probability of $50\%$ to
be assigned to the left or the right daughter node. In cases where all $k$ are
allocated to the same terminal node, the assignment of one category is flipped
to the other terminal node. If this procedure results in a non-valid split rule,
the two steps of random split variable selection and split point selection are
repeated. With the definition of $r=1$ and the selection of $v_1$ and $s_1$, the
initialization is complete and each individual of the population of trees is of
type $\tilde{\theta}_{1}= (v_1, s_1)$.

\subsection{Parent selection}
\label{parent}

In every iteration, each tree is selected once to be modified by one of the
variation operators. In cases where the crossover operator is applied, the
second parent is selected randomly from the remaining population. In this way,
some trees are selected more than once in each iteration.

\subsection{Variation operators}
\label{vari}

Four types of mutation operators and one crossover operator are utilized by our
algorithm. In each modification step, one of the variation operators is randomly
selected for each tree. The mutation and crossover operators are described
below.

\subsubsection{Split}

\textit{Split} selects a random terminal-node and assigns a valid, randomly
generated, split rule to it. As a
consequence, the selected terminal node becomes an internal node $r$ and
two new terminal nodes are generated.

The search for a valid split rule is conducted as in see Section~\ref{init}
for a maximum of $P$ iterations. 
In cases where no valid split rule can be assigned to the internal node at
position $r$, the search for a valid split rule is carried out on another
randomly selected terminal node. If, after 10~attempts, no valid split rule
can be found, then $\tilde{\theta}_{i+1} = \tilde{\theta}_{i}$. Otherwise, the
set of parameters in iteration $i+1$ are given by $\tilde{\theta}_{i+1}=
\tilde{\theta}_i \cup (v_{r},s_{r})$.

\subsubsection{Prune}

\textit{Prune} chooses a random internal node $r$, where $r>1$, which has two
terminal nodes as successors and prunes it into a terminal node. The tree's
parameters at iteration $i+1$ are $\tilde{\theta}_{i+1}=\tilde{\theta}_{i}
\setminus (v_{r},s_{r})$. If $\tilde{\theta}_{i}$ only comprises one internal
node, i.e., the root node, then $\tilde{\theta}_{i+1} = \tilde{\theta}_{i}$.

\subsubsection{Major split rule mutation}

\textit{Major split rule mutation} selects a random internal node $r$ and
changes the split rule, defined by the corresponding split variable $v_r$, and
the split point $s_r$. With a probability of $50\%$, a value from the range
$1,...,P$ is assigned to $v_r$. Otherwise $v_r$ remains unchanged and only $s_r$
is modified. Again, depending on the domain of $X_{v_r}$, either a random split
point from the range of possible values of $X_{v_r}$ is selected, or a non-empty
set of categories is assigned to each of the two terminal nodes. If the split
rule at $r$ becomes invalid, the mutation operation is reversed and the
procedure, starting with the selection of $r$, is repeated for a maximum of
3~attempts. Subsequent nodes that become invalid are pruned.

If no pruning occurs, $\tilde{\theta}_{i}$ and $\tilde{\theta}_{i+1}$ contain
the same set of parameters. Otherwise, the set of parameters $(v_{m_1}, s_{m_1},
..., v_{m_f}, s_{m_f})$, corresponding to invalid nodes, is removed from
$\tilde{\theta}_i$. Thus, $\tilde{\theta}_{i+1} = \tilde{\theta}_{i} \setminus
(v_{m_1}, s_{m_1}, ..., v_{m_f}, s_{m_f})$.

\subsubsection{Minor split rule mutation}

\textit{Minor split rule mutation} is similar to the \textit{major split rule
mutation} operator, but it does not alter $v_r$ and only changes the split point
$s_r$ by a minor degree. If $X_{v_r}$ is numerical or ordinal the split point
$s_r$ is changed by a non-zero number of unique values of $X_{v_r}$. In cases
where $X_{v_r}$ has less then $20$ unique values, the split point is change to
to the next larger, or the next lower, unique value of $X_{v_r}$. Otherwise,
$s_r$ is randomly shifted by a number of unique values that is not larger than
$10\%$ of the range of unique values of $X_{v_r}$. If $X_{v_r}$ is a nominal
variable, with less than $20$ categories, one of the categories is randomly
modified. Otherwise, at least one and at most $10\%$ of the variable's
categories are changed. In cases where subsequent nodes become invalid, further
split points are searched that preserve the tree's topology. After five
non-successful attempts at finding a topology preserving split point, the
non-valid nodes are pruned.

Equivalently to the \textit{major split rule mutation} operator the subsequent
solution $\tilde{\theta}_{i+1} = \tilde{\theta}_{i} \setminus (v_{m_1}, s_{m_1},
..., v_{m_f}, s_{m_f})$.

\subsubsection{Crossover}

\fixme{I really had problems following this section. The sub-operator is useful
but is it possible to introduce this with fewer sub- and super-scripts? It makes
it very challenging to read.}

\textit{Crossover} exchanges, randomly selected, subtrees between two trees. Let
$\tilde{\theta}_{i}^1$ and $\tilde{\theta}_{i}^2$ be the parameters of the trees
that are chosen for crossover. First, an internal node is selected randomly from
each of $\tilde{\theta}_{i}^1$ and $\tilde{\theta}_{i}^2$, which we denote as
$n_k^1$ and $n_k^2$. Let $\mathrm{sub}1(\tilde{\theta}_i^j, n^j_k)$ denote the subtrees
rooted by $n_k^j$ ($j=1,2$). Then, $\mathrm{sub}1(\tilde{\theta}_i^j) =
(v_{n^j_k}, s_{n^j_k},  v_{n^j_{k1}}, s_{n^j_{k1}}, \dots ,v_{n^j_{km}}, s_{n^j_{km}})$, where $(n^j_{k1}, ..., n^j_{km})$ are the set
of descendent node positions of $n^j_k$. Now, the two selected subtrees can be
written as $\mathrm{sub}1(\tilde{\theta}_{i}^1, n^1_k) $ and $\mathrm{sub}1(\tilde{\theta}_{i}^2,
n^2_k)$. Furthermore, $\mathrm{sub}2(\tilde{\theta}_{i}^1, n^1_k) = \tilde{\theta}_{i}^1
\setminus \mathrm{sub}1(\tilde{\theta}_{i}^1, n^1_k)$ and $\mathrm{sub}2(\tilde{\theta}_{i}^2,
n^2_k) = \tilde{\theta}_{i}^2 \setminus \mathrm{sub}1(\tilde{\theta}_{i}^2, n^2_k)$.

After the application of the \textit{crossover} operator, the trees' parameter
spaces are given by $\tilde{\theta}_{i+1}^1= \mathrm{sub}2(\tilde{\theta}_{i}^1, n^1_k)
\cup \mathrm{sub}1(\tilde{\theta}_{i}^2, n^2_k) \setminus (v_{m^2_{1}}, s_{m^2_{1}}, ...,
v_{m^2_{f}}, s_{m^2_{f}})$ and $\tilde{\theta}_{i+1}^2=
\mathrm{sub}2(\tilde{\theta}_{i}^2, n^2_k) \cup \mathrm{sub}1(\tilde{\theta}_{i}^1, n^1_k)
\setminus (v_{m^1_{1}}, s_{m^1_{1}}, ..., v_{m^1_{f}}, s_{m^1_{f}})$. Here,
$(v_{m^1_{1}}, s_{m^1_{1}}, ..., v_{m^1_{f}}, s_{m^1_{f}}) \subset
\mathrm{sub}1(\tilde{\theta}_i^1, n^1_k)$ and $(v_{m^2_{1}}, s_{m^2_{1}}, ...,
v_{m^2_{f}}, s_{m^2_{f}}) \subset \mathrm{sub}1(\tilde{\theta}_i^2, n^2_k)$ are the
parameters of the invalid nodes.

\subsection{Evaluation function}
\label{eval}

The evaluation function represents the requirements the population should adapt
to. In general, these requirements are formulated by Equation~\ref{eq:globally}.
A suitable evaluation function for classification and regression trees minimizes
the models' accuracy on the training data, and the models' complexity. This
subsection describes the currently implemented choices of evaluation functions
for classification and for regression.

\subsubsection{Classification}

The quality of a classification tree is most commonly measured as a function of
its misclassification MC and the complexity of a tree by a function
of the number of its terminal nodes $M$. \code{evtree} uses $2 N \cdot
\mathrm{MC}(Y, f(X, \theta))$ as a loss function. The number of terminal nodes,
weighted by $\log N$ and a user-specified parameter $\alpha$, measures the
complexity of trees.
%
\begin{eqnarray}
  \mathrm{loss}(Y, f(X, \theta)) & = & 2 N \cdot \mathrm{MC}(Y, f(X, \theta)) \nonumber \\
                                 & = & 2 \cdot \sum_{n=1}^N I(Y_n \neq f(X_{\cdot n}, \theta)),  \label{eq:classification} \\
  \mathrm{comp}(\theta)          & = & \alpha \cdot M \cdot \log N. \nonumber
\end{eqnarray}
%
With these particular choices, Equation~\ref{eq:globally} seeks trees
$\hat{\theta}$ that minimize the misclassification loss at a BIC-type tradeoff
with the number of terminal nodes.

Other, existing and commonly used choices of evaluation functions include the
\textit{Bayesian information criterion} \citep[BIC, as in][]{gray2008cta} and
\textit{minimum description length} \citep[MDL, as in][]{quinlan1989inferring}.
For both evaluation functions deviance is used for accuracy estimation. Deviance
is usually preferred over the misclassification rate in recursive partitioning
methods, as it is more sensitive to changes in the node probabilities
\citep[pp.~308--310]{hastie}. However, this is not necessarily an advantage for
global tree building methods like evolutionary algorithms.

\subsubsection{Regression}

For regression trees, accuracy is usually measured by the mean squared error MSE.
Here, it is again coupled with a BIC-type complexity measure:

Using $N \cdot \log \mathrm{MSE}$ as a loss function and
$\alpha \cdot 4 \cdot (M+1) \cdot \log N$ as the complexity part, the general
formulation of the optimization problem in  can be
rewritten as:
%
\begin{eqnarray} 
  \mathrm{loss}(Y, f(X, \theta)) & = & N \log \mathrm{MSE}(Y, f(X, \theta)) \nonumber \\
                                 & = & N \log \left\{ \sum_{n=1}^N (Y_n - f(X_{\cdot n}, \theta)^2) \right\}, \label{eq:regression} \\
  \mathrm{comp}(\theta)          & = & \alpha \cdot 4 \cdot (M+1) \cdot \log N. \nonumber
\end{eqnarray}
%
Here, $M+1$ is the effective number of estimated parameters, taking into account
the estimates of a mean parameter in each of the terminal nodes and the constant
error variance term. With $\alpha = 0.25$ the criteria is, up to a constant,
equivalent to the BIC used by \cite{fan2005rta}. However, the effective number
of parameters estimated for is actually much higher than $M+1$ due to the
selection of parameters in the split variable and the selection of the variable
itself. It is however unclear how these should be counted
\citep[p.~222]{gray2008cta, ripley}. Therefore, a more conservative default
value of $\alpha=1$ is assumed.

\subsection{Survivor selection}
\label{evo}

The population size stays constant during the evolution and only a fixed subset
of the candidate solutions can be kept in memory. A common strategy is the $(\mu
+ \lambda)$ selection, where $\mu$ survivors for the next generation are
selected from the union of $\mu$ parents and $\lambda$ offsprings. An
alternative approach is the $(\mu,\lambda)$ strategy where $\mu$
survivors for the next generation are selected from $\lambda$ offsprings.

Our algorithm uses $(1+1)$ selection, where one parent solution
competes with one offspring for a place in the population. In the case of a
mutation operator either the solution before $\tilde{\theta}_i$ or after
modification $\tilde{\theta}_{i+1}$ is kept in memory. In the case of the
crossover operator, the initial solutions of $\tilde{\theta}_i^1$ competes with
its subsequent solutions $\tilde{\theta}_{i+1}^1$. Correspondingly, one of the
two solutions $\tilde{\theta}_i^2$ and $\tilde{\theta}_{i+1}^2$ is rejected. The
survivor selection is done deterministically. The tree with lower fitness,
according to the evaluation function, is rejected. Note that, due to the
definition of the crossover operator, some trees are selected more than
once in an iteration. Correspondingly, these trees undergo the survival
selection process more than once in an iteration.

As in classification and regression tree analysis the individual solutions are
represented by trees. This design offers computational advantages over $(\mu +
\lambda)$, with $\mu>1$ and $\lambda>1$, and $(\mu,\lambda)$ strategies. In
particular, for the application of mutation operators no new trees have to be
constructed. The tree after modification is simply accepted or reversed to the
previous solution.

There are two important issues in the evolution process of an evolutionary
algorithm: population diversity and selective pressure
\citep{michalewicz1994genetic}. These factors are related, as with increasing
selective pressure the search is focused more around the currently best
solutions. An overly strong selective pressure can cause the algorithm to
converge early in local optima. On the other hand, an overly weak selective
pressure can make the search ineffective. Using a $(\mu + \lambda)$ strategy, a
strong selective pressure can occur in situations as follows. Suppose the $b$-th
tree of the population is one of the fittest trees in iteration $i$, and in
iteration $i$  one split rule of the $b$-th tree is changed only by a minor
degree. Then very few instances are classified differently and the overall
misclassification might not even change. However, as they represent one of the
best solutions in iteration $i$, they are both selected for the subsequent
population. This situation can occur frequently, especially when a fine-tuning
operator like the \textit{minor split rule mutation} is used. Then, the
diversity of different trees is lost quickly and the algorithm likely terminates
in a local optimum. The $(1+1)$ selection mechanism clearly avoids these
situations, as only the parent or the offspring can be part of the subsequent
population.

\subsection{Termination}
\label{term}

Using the default parameters,
the algorithm terminates when the quality of the best $5\%$ of trees stabilizes
for $100$ iterations, but not before $1000$ iterations. If the run does not
converge the algorithm terminates after a user-specified number of iterations.
In cases where the algorithm does not converge, a warning message is written to
the command line. The tree with the highest quality according to the evaluation
function is returned.

\section{Implementation and application in practice}
\label{impl}

Package \pkg{evtree} provides an efficient implementation of an evolutionary
algorithm that builds classification trees in \proglang{R}. CPU- and memory-
intensive tasks are fully computed in \proglang{C++}, while the user interfaces
and plot functions are written in \proglang{R}. The \code{.C()} interface
\citep{chambers2008software} was used to pass arguments between the two
languages. \pkg{evtree} depends on the \pkg{partykit} package 
\citep{hothorn2009partykit}, which provides an infrastructure for representing,
summarizing, and visualizing tree-structured models.

\subsection{User interface}

The principal function of the \pkg{evtree} package is the eponymous function
\code{evtree()} taking arguments
%
\begin{Code}
  evtree(formula, data = list(), weights = NULL, subset = NULL,
    control = evtree.control(...), ...)
\end{Code}
%
where \code{formula}, \code{data}, \code{weights}, and \code{subset} specify
the data in the usual way, e.g., via \code{formula = y ~ x1 + x2}.
Additionally, \code{control} comprises a list of control parameters
%
\begin{Code}
  evtree.control(minbucket = 7L, minsplit = 20L, maxdepth = 9L, 
    niterations = 10000L, ntrees = 100L, alpha = 1,
    operatorprob = list(pmutatemajor = 0.2, pmutateminor = 0.2,
      pcrossover = 0.2, psplit = 0.2, pprune = 0.2),
    seed = NULL, ...)
\end{Code}
%
where the parameters \code{minbucket}, \code{minsplit}, and \code{maxdepth}
constrain the solution to a minimum number of observations in each terminal
node, a minimum number of observation in each internal node and a maximum tree
depth. Note that the memory requirements increase by the square of the maximum
tree depth. Parameter \code{alpha} regulates the complexity parameter
$\alpha$ in Equation~\ref{eq:classification} and~\ref{eq:regression}, respectively.
\code{niterations} and \code{ntrees} specify the maximum number of iterations
and the number of trees in the population, respectively. With the argument
\code{operatorprob}, user-specified probabilities for the variation operators
can be defined. For making computations reproducible, argument \code{seed} is
an optional integer seed for the random number generator (at \proglang{C++} level).
If not specified, the random number generator is initialized by \verb|as.integer(runif(1, max = 2^16))|
in order to inherit the state of \code{.Random.seed} (at \proglang{R} level).
If set to \code{-1L}, the seed is initializied by the system time.

The trees computed by \code{evtree} inherit from class `\code{party}' supplied
by the \pkg{partykit} package. The methods inherited in this way
include standard \code{print()}, \code{summary()}, and \code{plot()} functions
to display trees and a \code{predict()} function to compute the fitted response
or node number etc.

\subsection{Case study: Customer targeting}

An interesting application for classification tree analysis is target marketing,
where limited resources are aimed at a distinct group of potential customers. An
example is provided by \cite{lilien2004} in the \textit{Bookbinder's Book Club}
marketing case study about a (fictitious) American book club. In this case study, a brochure of the book
``The Art History of Florence'' was sent to 20,000~customers, 1,806~of
which bought the book. The dataset contains a subsample of 1,300~customers for
building a predictive model of customer choice.

Besides predictive accuracy,
model complexity is a crucial issue in this application: Smaller trees are easier to
interpret and communicable to marketing experts and management professionals. 
Hence, we use \code{evtree} with a maximal depth of two levels of splits only.
This is contrasted with \code{rpart} and \code{ctree} with and without such a restriction
of tree depth to show that the evolutionary search of the global parameter space
can be much more effective in balancing predictive accuracy and complexity
compared to forward-search recursive partitioning.

All trees are constrained to have a minimum number of 10~observations per terminal node.
Additionally, a significance level of $1\%$ is employed in the construction of
conditional inference trees which is more appropriate than the default 5\%~level for
$1,300$~observations. To provide uniform visualizations and predictions of the
fitted models, `\code{party}' objects are used to represent all trees. For, `\code{rpart}'
trees \pkg{partykit} provides a suitable \code{as.party()} method while a
reimplementation of \code{ctree()} is provided in \pkg{partykit} (as opposed to
the original in \pkg{party}) that directly leverages the `\code{party}' infrastructure.

First, the data is loaded and the forward-search trees are grown
with and without depth restriction, visualizing the unrestricted trees
in Figure~\ref{fig:example_rpart_ctree}.
%
<<BBBClub-rpart-ctree, echo=TRUE, eval=FALSE>>=
data("BBBClub", package = "evtree")
library("rpart")
rp  <- as.party(rpart(choice ~ ., data = BBBClub, minbucket = 10))
rp2 <- as.party(rpart(choice ~ ., data = BBBClub, minbucket = 10,
  maxdepth = 2))
ct  <- ctree(choice ~ ., data = BBBClub, minbucket = 10, mincrit = 0.99)
ct2 <- ctree(choice ~ ., data = BBBClub, minbucket = 10, mincrit = 0.99,
  maxdepth = 2)
plot(rp)
plot(ct)
@
%
With the objective of building a smaller, but at still accurate tree,
\code{evtree} is constrained to a maximum tree depth of $2$,
see Figure~\ref{fig:example_evtree}.
%
<<BBBClub-evtree, echo=TRUE, eval=FALSE>>=
set.seed(1090)
ev <- evtree(choice ~ ., data = BBBClub, minbucket = 10, maxdepth = 2)
@
%
<<BBBClub-cache, echo=FALSE, eval=TRUE>>=
if(cache & file.exists("BBBClub-trees.rda")) {
load("BBBClub-trees.rda")
} else {
<<BBBClub-rpart-ctree>>
<<BBBClub-evtree>>
if(cache) {
  save(rp, rp2, ct, ct2, ev, file = "BBBClub-trees.rda")
} else {
  if(file.exists("BBBClub-trees.rda")) file.remove("BBBClub-trees.rda")
}
}
@
%
\setkeys{Gin}{width=1\textwidth}
\begin{figure}[p!]
\centering
\begin{minipage}{1\textwidth}
\begin{flushleft} \code{rpart} \end{flushleft}
\centering
\setkeys{Gin}{width=0.75\textwidth}
<<BBBClub-rpart-plot, fig=TRUE, echo=FALSE, width=8.1, height=7>>=
plot(rp)
@
\end{minipage}
\begin{minipage}{1\textwidth}
\begin{flushleft} \code{ctree} \end{flushleft}
\centering
\setkeys{Gin}{width=1\textwidth}
<<BBBClub-ctree-plot, fig=TRUE, echo=FALSE,  width=11.4, height=5.9>>=
plot(ct)
@
\end{minipage}
\caption{\label{fig:example_rpart_ctree} Trees for customer targeting constructed by \code{rpart}
  (upper panel) and \code{ctree} (lower panel). The target variable is the customer's choice of buying the book.
  The variables used for splitting are the number of art books purchased previously (\code{art}), the number
  of months since the first purchase (\code{first}), the frequency of previous purchases at the Bookbinder's Book Club
  (\code{freq}) and the customer's \code{gender}.}
\end{figure}
%
The resulting evolutionary tree is printed output below and visualized in
Figure~\ref{fig:example_evtree}.
%
<<BBBClub-evtree-display>>=
plot(ev)
ev
@
%
\setkeys{Gin}{width=1\textwidth}
\begin{figure}[t!]
\begin{flushleft} \code{evtree} \end{flushleft}
\centering
\setkeys{Gin}{width=0.7\textwidth}
<<BBBClub-evtree-plot, fig=TRUE, echo=FALSE, width=7.5, height=4.9>>=
plot(ev)
@
\caption{\label{fig:example_evtree} Tree for customer targeting constructed by \code{evtree}
   The target variable is the customer's choice of buying the book. The variables used for splitting are
   the number of art books purchased previously (\code{art}) and the number of months since the first purchase
    \code{first}.}
\end{figure}

Not surprisingly, the explanatory variable \code{art} -- the number of art books purchased previously at the
book club -- plays a key role in all constructed classification trees along with the number
of months since the first purchase (\code{first}), the frequency of previous purchases
(\code{freq}) and the customer's \code{gender}. Interestingly, though, the forward-search trees select
the arguably most important variable in the first split while the evolutionary tree uses \code{first}
in the first split and \code{art} in both second splits. Thus, the evolutionary tree
uses a different cutoff in \code{art} for bookclub members that joined in the last year as opposed
to older customers. While the former are predicted to be buyers if they previously bought
at least one art book, the latter are predicted to purchase the advertised art book only if they
previously bought at least two other art books. Certainly, this classification is easy to understand
and communicate (helped by Figure~\ref{fig:example_evtree}) to practitioners.

However, we still need to answer the question how well it performs in contrast to the other trees.
Hence, we set up a function \code{mc()} the computes the misclassification rate as a measure of
predictive accuracy and a function \code{evalfun()} that computes the evaluation function
(i.e., penalized by tree complexity) from Equation~\ref{eq:classification}.
%
<<evtree-performance>>=
mc <- function(obj) 1 - mean(predict(obj) == BBBClub$choice)
evalfun <- function(obj) 2 * nrow(BBBClub) * mc(obj) +
  width(obj) * log(nrow(BBBClub))
trees <- list("evtree" = ev, "rpart" = rp, "ctree" = ct, "rpart2" = rp2,
  "ctree2" = ct2)
round(sapply(trees, function(obj) c("misclassification" = mc(obj),
  "evaluation function" = evalfun(obj))), digits = 3)
@
%
Not surprisingly the evolutionary tree \code{ev} outperforms the depth-restricted
trees \code{rp2} and \code{ct2}, both in terms of misclassification and the penalized
evaluation function. However, it is interesting to see that \code{ev} performs even better
than the unrestricted conditional inference tree \code{ct} and is comparable in performance
to the unrestricted CART tree \code{rp}. Hence, the practitioner may choose the evolutionary
tree \code{ev} as it is the easiest to communicate.

Although the constructed trees are considerably different, the code above shows
that the predictive accuracy is rather similar. Moreover, below we see that the
structure of the individual predictions on the dataset are rather similar as well:
%
<<evtree-structure>>=
ftable(tab <- table(evtree = predict(ev), rpart  = predict(rp),
  ctree  = predict(ct), observed = BBBClub$choice))
sapply(c("evtree", "rpart", "ctree"), function(nam) {
  mt <- margin.table(tab, c(match(nam, names(dimnames(tab))), 4))
  c(abs = as.vector(rowSums(mt))[2],
    rel = round(100 * prop.table(mt, 1)[2, 2], digits = 3))
})
@
%
In this case, \code{evtree} classifies less
customers ($186$) as buyers as \code{rpart} ($216$) and \code{ctree} ($238$). However,
\code{evtree} achieves the highest proportion of correct classification among
the declared buyers: $72.6\%$ compared to $70.8\%$
(\code{rpart}) and $66.4\%$ (\code{ctree}).

In summary, this illustrates how \code{evtree} can be employed to better balance
predictive accuracy and complexity by searching a larger space of potential trees.
As a final note, it is worth pointing out that in this setup, several runs of
\code{evtree()} with the same parameters typically lead to the same tree. However,
this may not always be the case. Due to the stochastic nature of the search algorithm and
the vast search space, trees with very different structures but similar evaluation
function values may be found by subsequent runs of \code{evtree()}. Here, this
problem is alleviated by restricting the maximal depth of the tree, yielding a
clear solution.

\section{Performance comparison}
\label{comp}

In this section, we compare \code{evtree} with \code{rpart} and \code{ctree}
in a more rigorous benchmark comparison.

In the first part of the analysis (Section~\ref{comp1}) the algorithms are
compared on 14~benchmark datasets that are publicly available and $3$ real-world
datasets from the Austrian \textit{Diagnosis Related Group (DRG)} system
\citep{LKF2009}. The analysis is based on the evaluation of 250~bootstrap
samples for each of the 20~datasets. The misclassification rate on the
\textit{out-of-bag} \citep{hothorn2005design} samples is used as a measure of
predictive accuracy. Furthermore, the complexity is estimated by the number of
terminal nodes.

In the second part (Section~\ref{comp2}) the algorithms' performances are
assessed on an artificial chessboard problem that is simulated with different noise levels.
The estimation of predictive accuracy and the number of terminal nodes is based
on $250$ realizations for each simulation.

All models are constrained to a minimum number of 7~observations per terminal node,
20~observations per internal node and a maximum tree depth of~9. Apart from
that, the default settings of the algorithms are used. For assessment
of significant differences in predictive accuracy and complexity, respectively,
Dunnett's correction from \proglang{R}~package \pkg{multcomp} \citep{multcomp} was used
for calculating simultaneous $95\%$ confidence intervals on the individual datasets.

As missing values are currently not supported by \code{evtree} (e.g., by surrogate splits),
the 16~missing values in the \textit{Breast Cancer Database} -- the only dataset in the
study with missing values -- were removed before analysis.

\subsection{Benchmark and real-world problems}
\label{comp1}

In Table~\ref{tab:dataDescription} the benchmark and real-world datasets from
the Austrian DRG-system are described. 
\fixme{The DRG description is far too long. All readers are waiting for the results
of the benchmark now. One explanatory sentence -- or maybe two at most -- should be
used to state the general problem and point to the manual pages.}
In the Austrian DRG-system, resources are
allocated to hospitals by simple rules regarding diagnosis, selected individual
services, and patient characteristics. Within this system regression tree
analysis is performed to classify hospital patients into clinically meaningful
and comprehensible groups that consume similar hospital resources. The datasets
MEL0101 (intervention on the skull), HDG0202 (malignant neoplasms), and HDG0502
(acute affections of the respiratory tract and middle atelectasis) contain
patient groups with the same main diagnosis or procedure. The task of the
regression tree analysis is to further divide these groups with the intention of
finding groups with similar resource consumption. Possible split-variables are
the patients' diagnoses, procedures, age, and gender. The dependent variable is
the patients' length of stay, which has a good relationship with the total costs
and is available \citep{theurl2007ihf}. Permission to use the data was granted
by the \citet{BMG2009}. A more detailed description of the application can be
found in \citet{grubinger2010regression}.

\begin{table}[t!]
\centering
\begin{tabular}{lrrrrrr}
\toprule
{Dataset} & {Instances} &  \multicolumn{5}{l}{{Attributes}}   \\ 
           &             &  Binary              & Nominal          &  Ordered      &  Metric     & Classes \\ 
\midrule
Glass identification$^\#$  &   214   &  - &  - &  - & 9 &  6      	 \\
Statlog heart*         &   270   &  3 &   3 & 1 & 6 &   2     	 \\
Ionosphere$^\#$ &   351   &  2 &  - &  - & 32 & 2  	             \\
Musk$^+$  &  476  & - &  - &  - & 166 & 2                     	 \\
Breast cancer database$^\#$ &   685   &  - &  4 & 5 &  - &  2  	 \\
Pima Indians diabetes$^\#$ &   768   &  - &   - & - & 8 &  2      	 \\
Vowel$^\#$  &  990  & - &  1 &  - & 9 & 11     	                     \\
Statlog German credit* &  1000   &  2 &  10 & 1 & 7 &  2      	 \\
Contraceptive method* &   1437   & 3 &  - &  4 &  2 &  3  	     \\
DNA$^\#$  &  3186  & 180 &  - &  - & - & 3   	                       \\
Spam$^+$  &  4601  & - &  - &  - & 57 & 2                      	 \\
MAGIC gamma telescope*  &  19020  & - &  - &  - & 10 & 2     	   \\
Servo$^\#$ & 167 &  - & 4 & - &  -  &  -											   \\
Boston housing$^\#$ & 506 & 1 & - & - &  12  & -                 \\
MEL0101$^\diamond$ &   875 &  1 & 4 & 1 &  108  & -              \\
HDG0202$^\diamond$ &  3933 &  1 & 7 & 1 &   46  & -              \\
HDG0502$^\diamond$ & 8251 &  1 & 7 & 1 &   91  & -               \\
\bottomrule
\end{tabular}
\caption{\label{tab:dataDescription} Description of the evaluated benchmark datasets. The datasets marked with
  $*$ originate from the UCI machine learning repository \citep{Frank+Asuncion:2010} and are made available in the \pkg{evtree} package.
  Datasets marked with $\#$ and $+$ are from the \proglang{R} packages \pkg{mlbench} \citep{mlbench} and
  \pkg{kernlab} \citep{kernlab}, respectively. The three real-world datasets from the Austrian DRG system are marked with $\diamond$.} 
\end{table}

The relative performance of \code{evtree} and \code{rpart} is summarized in Figure~\ref{fig:benchmark} (upper panels). Performance differences are displayed relative to \code{evtree}'s performance. For example, on the \textit{Glass} dataset, the average misclassification rate of \code{rpart} is $2.7\%$ higher than the misclassification rate of \code{evtree}. It can be observed that on $12$ out of $17$ datasets \code{evtree} significantly outperforms \code{rpart} in terms of predictive accuracy. Only on the \textit{Contraceptive Method} dataset \code{evtree} performs slightly worse.  In terms of complexity, \code{evtree} models are significantly more complex on $10$ and less complex on $7$ datasets.

Figure~\ref{fig:benchmark} (lower panels) summarizes the relative performance of \code{evtree} and \code{ctree}. For 15 out of 17 datasets \code{evtree} shows a better predictive performance. The algorithms' performances  is significantly worse on the \textit{MEL0101} dataset, where the average misclassifiation rate of \code{ctree} is $5.6\%$ lower. However, on this dataset, \code{ctree} models are on average $86.5\%$ larger than \code{evtree} models. The relative complexity of \code{evtree} models is significantly smaller for 15 and larger for 1 dataset.

<<benchmark-results, echo=FALSE>>=
## load results
for(i in Sys.glob("results/*.RData")) load(i)

## preprocess for reference evtree
preprocess <- function(d, dname = "datasetname", isclassification = TRUE){
    if(isclassification) d[, 1:3] <- 1 - d[ ,1:3]
    d <- as.data.frame(d)
    colnames(d) <- c("evtree", "rpart", "ctree","evtree", "rpart", "ctree")
    for(i in 3:1) d[, i] <- d[, i] / d[, 1] * 100
    for(i in 6:4) d[, i] <- d[, i] / d[, 4] * 100
    x <- d[, 1:3]
    y <- d[, 4:6]
    rval <- reshape(x, idvar="samp", times=names(x), timevar = "alg",varying= list(names(x)), direction="long")
    names(rval)[2] <- "accuracy"
    rval$complexity <- reshape(y, idvar="samp", times=names(y), timevar = "alg",varying= list(names(y)), direction="long")[,2]
    rval$alg <- factor(rval$alg, levels = c("evtree", "ctree", "rpart"))
    rval$ds <- dname
    rval
}

## collect results for all datasets
r <- rbind(
  preprocess(d = rglass, dname = "Glass identification", isclassification = TRUE),
  preprocess(d = rheart, dname = "Statlog heart", isclassification = TRUE),
  preprocess(d = rionosphere, dname = "Ionosphere", isclassification = TRUE),
  preprocess(d = rmusk, dname = "Musk", isclassification = TRUE),
  preprocess(d = rbreastcancer, dname = "Breast cancer database", isclassification = TRUE),
  preprocess(d = rpima, dname = "Pima Indians diabetes", isclassification = TRUE),
  preprocess(d = rvowel, dname = "Vowel", isclassification = TRUE),
  preprocess(d = rcredit, dname = "Statlog German credit", isclassification = TRUE),
  preprocess(d = rcontraceptive, dname = "Contraceptive method", isclassification = TRUE),
  preprocess(d = rdna, dname = "DNA", isclassification = TRUE),
  preprocess(d = rspam, dname = "Spam", isclassification = TRUE),
  preprocess(d = rmagicgamma, dname = "Magic gamma telescope", isclassification = TRUE),
  preprocess(d = rservo, dname = "Servo", isclassification = FALSE),
  preprocess(d = rbostonhousing, dname = "Boston housing", isclassification = FALSE),
  preprocess(d = rmel0101, dname = "MEL0101", isclassification = FALSE),
  preprocess(d = rhdg0202, dname = "HDG0202", isclassification = FALSE),
  preprocess(d = rhdg0502, dname = "HDG0502", isclassification = FALSE)
)
r$ds <- factor(r$ds)
r$samp <- factor(r$samp)
r$dssamp <- r$ds:r$samp

## compute multiple comparisons
library("multcomp")
cstats <- function(alg = "rpart", value = "accuracy", data = r) {
  dlab <- rev(unique(data$ds))
  k <- length(dlab)
  mean  <- numeric(k)
  lower <- numeric(k)
  upper <- numeric(k)
  names(data)[names(data) == value] <- "value"
  for(i in 1:k) {
    mod1 <- lm(value ~ alg, data = subset(data, ds == dlab[i]))
    pt <- glht(mod1, linfct = mcp(alg = "Dunnett"))
    w <- confint(pt)$confint
    d <- which(levels(r$alg) == alg) - 2
    mean[i]  <-  w[1+d]
    lower[i] <-  w[3+d]
    upper[i] <-  w[5+d]
  }
  rval <- data.frame(mean, lower, upper)
  rownames(rval) <- dlab
  return(rval)
}

acc_rpart <- cstats("rpart", "accuracy")
com_rpart <- cstats("rpart", "complexity")
acc_ctree <- cstats("ctree", "accuracy")
com_ctree <- cstats("ctree", "complexity")

## function for visualization
ciplot <- function(x, xlim = NULL, main = "", xlab = "", ylab = TRUE) {
  nam <- rownames(x)
  k <- length(nam)
  plot(x$mean, 1:k, xlim = xlim, axes = FALSE, xlab = "", ylab = "", pch = 19)
  arrows(x$lower, 1:k, x$upper, 1:k, angle = 90, length = 0.05, code = 3)
  if(xlab == "") axis(1, labels = FALSE) else axis(1)
  if(ylab) ylab <- nam
  axis(2, at = 1:k, labels = ylab, las = 1, cex = 0.8)  
  axis(2, at = k + 1.5, labels = main, tick = FALSE, las = 1, outer = TRUE, cex.axis = 1.5, xpd = TRUE)
  mtext(xlab, side = 1, line = 3, xpd = TRUE)
  abline(h = 5.5)
  abline(v = 0, lty = 2)  
  box()
}
@

\setkeys{Gin}{width=1\textwidth}
\begin{figure}[t!]
\centering
<<benchmark-plot, fig = TRUE, echo = FALSE, width=10, height=10>>=
par(mfrow = c(2, 2), oma = c(5, 10, 2, 0), mar = c(1, 1, 2, 1))

xlim1 <- range(cbind(acc_rpart, acc_ctree))
xlim2 <- range(cbind(com_rpart, com_ctree))

ciplot(acc_rpart, xlim = xlim1, main = "rpart", ylab = TRUE, xlab = "")
ciplot(com_rpart, xlim = xlim2, main = "",      ylab = FALSE, xlab = "")
ciplot(acc_ctree, xlim = xlim1, main = "ctree", ylab = TRUE,
  xlab = "relative difference in predictive accuracy (%)")
ciplot(com_ctree, xlim = xlim2, main = "",      ylab = FALSE,
  xlab = "relative difference in complexity (%)")
@
\caption{\label{fig:benchmark} Performance comparison of \code{evtree} vs.\ \code{rpart} (upper panels) and \code{evtree} vs.\ \code{ctree} (lower panels).
  Prediction error (left panels) is compared by the relative difference of the misclassification rate or the mean-squared error.
  The complexity (right panels) is compared by the relative difference of the number of terminal nodes.}
\end{figure}

The disadvantages of the \code{evtree} algorithm are computation time and memory requirements. While the smallest of the analyzed datasets, \textit{Glass identification},
only needed approximately 4--6~seconds to fit, larger datasets demanded several minutes. The fit of a model from the largest dataset, \textit{MAGIC gamma telescope}, required approximately 40--50 minutes and a main memory of 400 Mbit. The required resources were measured on an Intel Core 2 Duo with 2.2 GHz and 2 GB RAM using the 64-bit version of Ubuntu 10.10.
\fixme{also briefly discuss the issue of instability of the evolutionary trees even for large populations and number of runs}

\subsection{Artificial problem}
\label{comp2}

In this section we demonstrate the ability of \code{evtree} to solve an artificial problem that is difficult to solve for most recursive classification tree algorithms \citep{loh2009improving}. The data was simulated with $2000$ instances for both the training-set and the test-set. Predictor variables $X_1$ and $X_2$ are simulated to be uniformly distributed in the interval $[0,4]$. The classes are distributed in alternating squares forming a $4 \times 4$ chessboard in the $(X_1, X_2)$-plane. One realization of the simulated data is shown in Figure~\ref{fig:chess}. Furthermore, variables $X_3-X_8$ are noise variables that are uniformly distributed on the interval [0, 1]. The ideal model for this problem only uses variables $X_1$ and $X_2$ and has 16 terminal nodes, whereas each terminal node comprises the observations that are in the region of one square. Two further simulations are done in the same way, but $5\%$ and $10\%$ percent of the class labels are randomly changed to the other class.

<<chessboard, echo=FALSE>>=
chessboard44 <- function(n = 4000, noisevariables = 6, noise = 0) {
  chess44 <- array(0,c(n,noisevariables+3))
  for(i in 1:(noisevariables+2))
      chess44[,i] <- as.numeric(runif(dim(chess44)[1]))*4

   x <- chess44[,1]
   y <- chess44[,2]
   chess44[, ncol(chess44)] <- 0
   for(k in 1:4)  
      chess44[(x <= k & x > k-1 & y <= k & y > k-1), ncol(chess44)] <- 1
   for(k in 1:2)  
      chess44[(x <= k & x > k-1 & y <= k+2 & y > k+1), ncol(chess44)] <- 1
   for(k in 1:2)  
      chess44[(y <= k & y > k-1 & x <= k+2 & x > k+1), ncol(chess44)] <- 1

   if(noise > 0) {
      flipclasslist <- sample(n, n * (noise / 100), replace = FALSE)

      for(i in 1:length(flipclasslist)){
	  if(chess44[flipclasslist[i], ncol(chess44)] == 1)
	      chess44[flipclasslist[i], ncol(chess44)] = 0
	  else if(chess44[flipclasslist[i], ncol(chess44)] == 0)
	      chess44[flipclasslist[i], ncol(chess44)] = 1
      }
  }

  chess44 <- as.data.frame(chess44)
  chess44[,ncol(chess44)] <- as.factor(chess44[,ncol(chess44)])
  names(chess44) <- c(paste("X", 1:8, sep = ""), "Y")
  chess44
}
@

\setkeys{Gin}{width=0.5\textwidth}
\begin{figure}[t!]
\centering
<<chessboard-plot, height=4.5, width=4.5, fig = TRUE, echo = FALSE>>=
chess44 <- chessboard44(2000)
plot(X2 ~ X1, data = chess44, xlim = c(0, 4), ylim = c(0, 4), pch = c(1, 4)[Y], col = c("black", "slategray")[Y])
@
\caption{\label{fig:chess} Class distribution of the simulated $4 \times 4$ chessboard problem with zero noise,
  plotted on the $(X_1, X_2)$-plane. The two classes are indicated by black circles and gray crosses, respectively.}
\end{figure}

The results are summarized in Table~\ref{tab:resultsChessboard}. It can be seen that, in the absence of noise, \code{rpart} models on average classify $69.1\%$ of the data points correctly and had $16.6$ terminal nodes. An average \code{ctree} model only has $1.1$ terminal nodes and a classification accuracy of $49.9\%$. In contrast, \code{evtree} classifies $93.2\%$ of the instances correctly and requires $14.4$ terminal nodes. In the presence of $5\%$ and $10\%$ noise, \code{evtree} classifies $89.0\%$ and $84.5\%$ of the data correctly.

<<chessboard-table, echo=FALSE, results=tex>>=
library("xtable")
load("./results/chessboard44_0.RData")
load("./results/chessboard44_5.RData")
load("./results/chessboard44_10.RData")

chesstable_means  <- as.data.frame( rbind(apply(rchessboard44_0,2,mean), apply(rchessboard44_5,2,mean) , apply(rchessboard44_10,2,mean) )) 
names(chesstable_means) <-  c("\\code{evtree}", "\\code{rpart}", "\\code{ctree}", "\\code{evtree}", "\\code{rpart}", "\\code{ctree}")
chesstable_means[,1:3] <-  format(chesstable_means[,1:3]*100, digits=1, nsmall=1)
chesstable_means[,4:6] <-  format(chesstable_means[,4:6], digits=1, nsmall=1)

chesstable_sd  <- as.data.frame( rbind(apply(rchessboard44_0,2,sd), apply(rchessboard44_5,2,sd) , apply(rchessboard44_10,2,sd) )) 
names(chesstable_sd) <-  c("\\code{evtree}", "\\code{rpart}", "\\code{ctree}", "\\code{evtree}", "\\code{rpart}", "\\code{ctree}")
chesstable_sd[,1:3] <-  format(chesstable_sd[,1:3]*100, digits=1, nsmall=1)
chesstable_sd[,4:6] <-  format(chesstable_sd[,4:6], digits=1, nsmall=1)

chesstable <- chesstable_means
for(j in 1:ncol(chesstable_means)){
	for(i in 1:nrow(chesstable_means)){
		chesstable[i,j] <- paste(chesstable_means[i,j] ,  "(", chesstable_sd[i,j], ")",  sep="")	
	}
}

chesstable <- cbind(as.integer(rbind(0,5,10)), chesstable)
colnames(chesstable)[1] = ""

print(xtable(chesstable,
caption = "Mean (and standard deviation) of accuracy and number of terminal nodes for simulated $4 \\times 4$ chessboard examples.",
caption.placement= "bottom",
label= "tab:resultsChessboard"), 
include.rownames = FALSE, allign= "rllllll", hline.after=NULL,
sanitize.text.function = identity,
add.to.row=list(pos=list(-1,-1, 0, 3), command=c(
"\\toprule", 
c("\\multicolumn{1}{l}{Noise (\\%)} & \\multicolumn{3}{l}{Accuracy}  & \\multicolumn{3}{l}{Terminal nodes}\\\\",
"\\midrule",
"\\bottomrule"
)))
)
@

\section{Conclusions}
\label{conc}

In this paper, we presented the \pkg{evtree} package, which implements
classification and regression trees that are grown by through an evolutionary
algorithm. The package uses standard \code{print()}, \code{summary()}, and
\code{plot()} functions to display trees and a \code{predict()} function to
predict the class labels of new data from the \pkg{partykit} package. As evolutionary
learning of trees is computationally demanding, most calculations
are conducted in \proglang{C++}. At the moment our algorithm does not support
parallelism. However, we intend to extend \pkg{evtree} to parallel computing.

The comparisons with recursive partitioning methods \code{rpart}
and \code{ctree} in Sections~\ref{impl} and~\ref{comp} shows that \code{evtree}
performs very well in a wide variety of settings, often balancing predictive
accuracy and complexity better than the forward-search methods.

However, the goal of \code{evtree} is not to replace the well-established algorithms like
\code{rpart} and \code{ctree} but rather to complement the tree toolbox with
an alternative method which may perform better given sufficient amounts of time and main memory.
By the nature of the algorithm it is able to discover patterns which cannot be modeled by a greedy
forward-search algorithm. As \code{evtree} models can be substantially different to recursively
fitted models, it can be beneficial to use both approaches, as this may reveal
additional relationships in the data.

\bibliography{evtree}

\end{document}