### 2013-10-05 pkg/partykit/inst/mloss/mloss.Rnw
added old MLOSS paper (with some small touch-ups)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@351 255f4249-2263-4417-b1ea-8498779e27a7

\documentclass[a4paper]{article}
\usepackage{jmlr2e}
\usepackage{color,hyperref}

%% software markup
\let\proglang=\textsf
\let\pkg=\textit
\let\code=\texttt

%% \usepackage{Sweave} is essentially
\RequirePackage[T1]{fontenc}
\RequirePackage{ae,fancyvrb}
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
\newenvironment{Schunk}{}{}
\setkeys{Gin}{width=.7\textwidth}
\SweaveOpts{engine=R, eps=FALSE, keep.source = TRUE, results=hide}
<<options, echo=FALSE>>=
options(prompt = "R> ", continue = "+  ", digits = 4)
@

%% header declarations
\author{Torsten Hothorn \and Achim Zeileis \and Kurt Hornik \\
        \addr{\proglang{R} Project for Statistical Computing and Graphics} \\
	\addr{E-mail: Firstname.Lastname@R-project.org}}
\title{Let's Have a \pkg{party}!\\An Open-Source Toolbox for Recursive Partytioning}
\editor{(initial submission for MLOSS track)}

% hyperref setup
\definecolor{Red}{rgb}{0.5,0,0}
\definecolor{Blue}{rgb}{0,0,0.5}
\hypersetup{%
  pdftitle = {Let's Have a party! An Open-Source Toolbox for Recursive Partytioning},
  pdfsubject = {submitted to JMLR-MLOSS},
  pdfkeywords = {open-source software, recursive partitioning, decision trees, statistical learning},
  pdfauthor = {Torsten Hothorn, Achim Zeileis, Kurt Hornik},
  %% change colorlinks to false for pretty printing
  colorlinks = {true},
  linkcolor = {Blue},
  citecolor = {Blue},
  urlcolor = {Red},
  hyperindex = {true},
  linktocpage = {true},
}

\begin{document}

\maketitle

\begin{abstract}
Package \pkg{party}, implemented in the \proglang{R} system for statistical
computing, provides basic classes and methods for recursive partitioning
along with reference implementations for three recently suggested tree-based
learners: conditional inference trees and forests, and model-based
recursive partitioning.
\end{abstract}

\begin{keywords}
open-source software, recursive partitioning, decision trees, statistical learning. 
\end{keywords}

\section{Overview}

Tree-based learners, also known as recursive partitioning techniques, belong to the
core tools for classification and regression, both in the machine
learning and statistical learning communities. Among the most popular representatives
are C4.5 \citep{mloss:Quinlan:1993} and CART \citep{mloss:Breiman+Friedman+Olshen:1984},
respectively, which were both accompanied by non-free implementations of the algorithms.
Following the success of these seminal algorithms, many modifications, extensions,
and enhancements of the basic recursive partitioning idea have been proposed in 
the literature---more often than not with proprietary or closed-source implementations or
no software at all. A notable exception is the open-source \proglang{Java} package \pkg{Weka}
\citep{mloss:Witten+Frank:2005} that provides free de facto reference implementations
of tree-based learners (including C4.5) or original implementations of novel
algorithms (such as LMT). In statistical learning, the ``lingua franca'' is
\proglang{R} \citep{mloss:R:2007}, an open-source programming environment that 
also provides a reference implementation for CART in package \pkg{rpart}.

Here, we present the \proglang{R} package \pkg{party}, a toolbox for recursive
partytioning. Unlike the implementations above, it does not only implement
specific algorithms, but aims to provide basic infrastructure for tree-based
learners along with flexible and adjustable learning algorithms built on top.
The basic data structures encompass a general high-level class \code{BinaryTree} 
defined in \proglang{R} for
representing various types of binary trees where arbitrary types of information 
can be attached to the nodes. It comes with generic functionality for summarizing
such trees, creating visualizations and computing predictions. This generic
infrastructure has been employed for implementing three flexible recursive
partitioning algorithms. The first is the conditional inference tree framework of
\cite{mloss:Hothorn+Hornik+Zeileis:2006} which learns classical trees with
constant fits in the terminal nodes, based on permutation tests for split selection.
Its user interface is the function \code{ctree} that can be applied out of the box
to regression and classification problems where both inputs and outputs can be
categorical, numerical, censored, or multivariate. Re-using this basic learner,
the package also provides an ensemble method for fitting forests of trees
in the function \code{cforest}. A key advantage of this random forest type
algorithm is that it can produce unbiased variable importance measures
\citep{mloss:Strobl+Boulesteix+Zeileis:2007}. Finally, function \code{mob}
provides a model-based recursive partitioning learner \citep{mloss:Zeileis+Hothorn+Hornik:2008}
that allows to attach arbitrary parametric models
(such as, e.g., linear or logistic regression, multivariate or survival models)
to the nodes in the tree and uses structural change tests for split selection.

The package \pkg{party} and all the components it depends on---including packages
with tools for model construction, with visualization infrastructure and with
inference techniques---are available under the General
Public License (GPL) from the Comprehensive \proglang{R} Archive Network (CRAN) at
\url{http://CRAN.R-project.org/}. The package contains \proglang{R} and \proglang{C}
code and has been successfully built on a multitude of platforms including Windows,
Mac~OS and various flavors of Linux.  It can be installed automatically from within
\proglang{R} via the command \code{install.packages("party", dependencies = TRUE)}
(which also installs all packages needed for examples); to get started, see the
online reference manual, e.g., via \code{help(package = "party")}. More advanced
user- and developer-level documentation is available in the package vignettes (see
\code{vignette(package = "party")}) and the \proglang{C} source code \pkg{doxygen}
documentation (in the \code{inst/documentation} sub-directory of the package sources).

In the following, we give a brief tour of the package by applying the three core
algorithms to classification problems before discussing some details of the developer-level
features.

\section{User interface}

<<GlaucomaM, echo=FALSE>>=
library("partykit")
data("GlaucomaM", package = "TH.data")
set.seed(10131218)
idx <- sample(1:nrow(GlaucomaM), 150)
GM1 <- GlaucomaM[idx,]
GM2 <- GlaucomaM[-idx,]
@
%% geeky detail: random seed was chosen as
%% str2num <- function(x) as.numeric(paste(gsub(" ", "0",
%%   format(c(10, match(tolower(strsplit(x, "")[[1]]), letters)))[-1]), collapse = ""))
%% str2num("JMLR")

To demonstrate usage of the high-level algorithms, we present some short examples
for fitting single trees, random forests and model-based trees in \pkg{party}
for two different binary classification tasks.

\paragraph{Trees and forests.} The \code{GlaucomaM} data set contains \Sexpr{nrow(GlaucomaM)}
observations of glaucoma classification in human eyes (affected vs.\ healthy) 
along with \Sexpr{ncol(GlaucomaM)-1} features derived from confocal laser scanning images
of the optic nerve head. In the \proglang{R} session below, we first load
package and data and then split it randomly into training and test sets of
150 and \Sexpr{nrow(GlaucomaM) - 150} observations, respectively.
<<GlaucomaM1, eval=FALSE>>=
<<GlaucomaM>>
@
A \code{ctree} for predicting the \code{Class} variable, as explained by (\code{\~})
all other variables (\code{.}) contained in the learning set \code{GM1}, can be fitted via
<<ctree>>=
ct <- ctree(Class ~ ., data = GM1)
@
creating a fitted \code{BinaryTree} object \code{ct}. For obtaining a short summary
of the tree it can be either printed by typing \code{ct} at the command line (or more
explicitly: \code{print(ct)}) or visualized by \code{plot(ct)}. The resulting plot
(depicted in the left panel of Figure~\ref{fig}) shows that the glaucoma probability 
increases clearly for lower volumes of the optic nerve head 
(as captured in the different volume measures
\code{vari} and \code{vasg}). After three splits, no more significant splits were found.
To evaluate the performance of this fitted tree on the test data \code{GM2}, we can
compute the confusion matrix of true and predicted class memberships
<<ctree-confusion, results=verbatim>>=
table(true = GM2$Class, pred = predict(ct, newdata = GM2))
@
corresponding to a misclassification rate of
\Sexpr{100 * round(1 - sum(diag(table(true = GM2$Class, pred = predict(ct, newdata = GM2))))/nrow(GM2), digits = 3)}\%.
Of course, this is only a rough ad hoc estimate of the generalization error, for a more
thorough evaluation and comparison with other tree algorithms see the original publications
mentioned above.

\paragraph{Model trees.} For illustrating the \code{mob} algorithm, we partition a logistic
regression for \code{diabetes} explained by (\code{\~}) \code{glucose} in the
well-known Pima Indians Diabetes data set, using five remaining variables for partitioning.
<<PimaIndiansDiabetes>>=
data("PimaIndiansDiabetes", package = "mlbench")
PID <- subset(PimaIndiansDiabetes, glucose > 0 & pressure > 0 & mass > 0)
@
This loads the data set and selects a subset \code{PID} of \Sexpr{nrow(PID)} observations
excluding physically impossible zero measurements in some of the variables. The \code{mob}
based on a logistic regression (i.e., a generalized linear model with binomial family)
can be fitted via
<<mob>>=
mb <- glmtree(diabetes ~ glucose | pregnant + pressure + mass + pedigree + age,
  data = PID, family = binomial)
@
Its visualization produced by \code{plot(mb)} is depicted in the right panel of
Figure~\ref{fig}. It shows the empirical \code{diabetes} proportions over \code{glucose}
groups along with the predictions from the logistic regression, conveying that the overall
level (regression intercept) increases with body mass index and age while the glucose
effect (regression slope) decreases. For a benchmark comparison of this algorithm with
other trees (including \pkg{Weka}'s LMT and J4.8 implementation of C4.5) see
\cite{mloss:Zeileis+Hothorn+Hornik:2008}.

\setkeys{Gin}{width=\textwidth}
\begin{figure}[t!]
\begin{center}
<<mob-plot, echo=FALSE, fig=TRUE, height=6, width=14>>=
grid.newpage()
pushViewport(viewport(layout = grid.layout(1, 2)))
pushViewport(viewport(layout.pos.col = 1, layout.pos.row = 1))
plot(ct, newpage = FALSE, pop = TRUE,
  tp_args = list(reverse = FALSE, fill = grey(c(0.45, 0.8))))
popViewport()
pushViewport(viewport(layout.pos.col = 2, layout.pos.row = 1))
plot(mb, newpage = FALSE, pop = TRUE, tp_args = list(ylines = 1.3,
  margins = c(1.5, 1, 1.5, 2.2), gp = gpar(fill = grey(c(0.45, 0.8))),
  linecol = "black"))
popViewport(2)
@
\caption{\label{fig} Tree visualizations: distribution of glaucoma
for fitted \code{ctree} (left) and distribution of diabetes over glucose
for logistic-regression \code{mob} (right).}
\end{center}
\end{figure}

%% This very brief tour of the features in \pkg{party} only hinted at the
%% flexibility of the tools, some more information is given below. Detailed
%% information on arguments and control options is available in the
%% manual pages (\code{help("ctree", package = "party")} etc.) and
%% the package vignettes.

\section{Internal structure}

The \pkg{party} package contains \proglang{R} code defining classes and 
methods along with high-level algorithms built on top of these, as well as
an additional program layer in \proglang{C} that implements time-critical operations
by directly modifying previously created \proglang{R} objects.
In particular, the \proglang{R} building blocks encompass
classes and methods for \code{LearningSample} and \code{BinaryTree} objects
which can be re-used in high-level algorithms (such as
\code{ctree}, \code{cforest} and \code{mob}) or directly at \proglang{C} level.
\code{LearningSample} is a data set---created from external sources or conversion
from other objects---which is enhanced by meta-information such as variable
orderings that can speed up split searches. \code{BinaryTree} is
a tree representation based on a recursive list structure
containing node information (e.g., primary splits, surrogate splits, summary statistics,
predictions, etc.) as well as left and right daughter nodes. Recursive methods
for predicting future cases, textual and graphical representations etc., 
are defined for objects of class \code{BinaryTree}.

The tree-growing can be controlled in various ways: starting from simple 
hyper-parameters, such as minimal leaf size, significance level or 
maximal depth, to very complex arguments (e.g., parametric model classes 
for \code{mob} or visualization functions for terminal nodes). At the \proglang{R}
level, users and developers can make use of the documented API for
inspecting and manipulating \code{BinaryTree} objects and, in addition,
are free to extend currently implemented methodology.

Unit testing of \pkg{party} (as for more than 1,200 other \proglang{R} packages on
CRAN) is performed automatically on a daily basis for various platforms and
versions of \proglang{R}.

\section{Conclusions}

The package \pkg{party} extends the set of machine learning techniques
available in the \proglang{R} system for statistical computing
(see the corresponding CRAN task view at \url{http://CRAN.R-project.org/src/contrib/Views/})
with a toolbox for recursive partitioning. The package provides basic
infrastructure for binary trees along with reference implementations of 
three recently-suggested tree-based learners: conditional inference trees and forests,
and model-based recursive partitioning. Party on!

\bibliography{mloss}

\end{document}