### 2014-02-07 pkg/partykit/inst/Psychoco-2014/slides.Rnw
slides for Psychoco 2014
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@451 255f4249-2263-4417-b1ea-8498779e27a7

\documentclass[11pt,compress,t]{beamer}
\usetheme{Z}
\usepackage{amsfonts,amstext,amsmath}
%% need no \usepackage{Sweave}
\definecolor{InputColor}{rgb}{0,0,0.3}
\definecolor{OutputColor}{rgb}{0.2,0.2,0.2}

\graphicspath{{images/}}

%% commands
\newcommand{\ui}{\underline{i}}
\newcommand{\oi}{\overline{\imath}}
\newcommand{\darrow}{\stackrel{\mbox{\tiny \textnormal{d}}}{\longrightarrow}}
\newcommand{\parrow}{\stackrel{\mbox{\tiny \textnormal{p}}}{\longrightarrow}}
\newcommand{\dotequals}{\stackrel{\cdot}{=}}
\newcommand{\efp}{\mathrm{\it efp}}
\newcommand{\given}{\, | \,}
\newcommand{\ltime}{\lambda_\mathrm{time}}
\newcommand{\lcomp}{\lambda_\mathrm{comp}}
\newcommand{\argmin}{\operatorname{argmin}\displaylimits}

\SweaveOpts{engine=R, eps=FALSE, keep.source=TRUE}

<<packages, echo=FALSE, results=hide>>=
library("partykit")
library("psychotools")
data("Topmodel2007", package = "psychotree")
options(digits = 4)
@

<<btfit2, echo=FALSE, results=hide>>=
btfit2 <- function(y, x = NULL, start = NULL, weights = NULL,
  offset = NULL, ..., estfun = FALSE, object = FALSE) {
  rval <- btReg.fit(y, weights = weights, ...,
    estfun = estfun, vcov = object)
  list(
    coefficients = rval$coefficients,
    objfun = -rval$loglik,
    estfun = if(estfun) rval$estfun else NULL,
    object = if(object) rval else NULL
  )
}
@

<<estfun.btReg, echo=FALSE, results=hide>>=
estfun.btReg <- function(x, ...) x$estfun
@

<<panelfunction, echo=FALSE, results=hide>>=
## visualization function
node_btplot <- function(mobobj, id = TRUE,
  worth = TRUE, names = TRUE, abbreviate = TRUE, index = TRUE, ref = TRUE,
  col = "black", linecol = "lightgray", cex = 0.5, pch = 19, xscale = NULL, yscale = NULL, ylines = 1.5)
{
    ## node ids
    node <- nodeids(mobobj, terminal = FALSE)

    ## get all coefficients 
    cf <- partykit:::apply_to_models(mobobj, node, FUN = function(z)        
      if(worth) worth(z) else coef(z, all = FALSE, ref = TRUE))
    cf <- do.call("rbind", cf)
    rownames(cf) <- node

    ## get one full model
    mod <- partykit:::apply_to_models(mobobj, node = 1L, FUN = NULL)

    if(!worth) {
      if(is.character(ref) | is.numeric(ref)) {
        reflab <- ref
        ref <- TRUE
      } else {
        reflab <- mod$ref
      }
      if(is.character(reflab)) reflab <- match(reflab, mod$labels)
      cf <- cf - cf[,reflab]
    }

    ## reference
    if(worth) {
      cf_ref <- 1/ncol(cf)
    } else {
      cf_ref <- 0
    }

    ## labeling
    if(is.character(names)) {
      colnames(cf) <- names
      names <- TRUE
    }

    ## abbreviation
    if(is.logical(abbreviate)) {
      nlab <- max(nchar(colnames(cf)))
      abbreviate <- if(abbreviate) as.numeric(cut(nlab, c(-Inf, 1.5, 4.5, 7.5, Inf))) else nlab
    }
    colnames(cf) <- abbreviate(colnames(cf), abbreviate)

    if(index) {
      x <- 1:NCOL(cf)
      if(is.null(xscale)) xscale <- range(x) + c(-0.1, 0.1) * diff(range(x))
    } else {
      x <- rep(0, length(cf))
      if(is.null(xscale)) xscale <- c(-1, 1)      
    }
    if(is.null(yscale)) yscale <- range(cf) + c(-0.1, 0.1) * diff(range(cf))

    ## panel function for bt plots in nodes
    rval <- function(node) {

      ## node index
      id <- id_node(node)

      ## dependent variable setup
      cfi <- cf[id,]

      ## viewport setup
      top_vp <- viewport(layout = grid.layout(nrow = 2, ncol = 3,
    			 widths = unit(c(ylines, 1, 1), c("lines", "null", "lines")),  
        		 heights = unit(c(1, 1), c("lines", "null"))),
    			 width = unit(1, "npc"), 
    			 height = unit(1, "npc") - unit(2, "lines"),
        		 name = paste("node_btplot", id, sep = ""))
      pushViewport(top_vp)
      grid.rect(gp = gpar(fill = "white", col = 0))

      ## main title
      top <- viewport(layout.pos.col = 2, layout.pos.row = 1)
      pushViewport(top)
      mainlab <- paste(ifelse(id, paste("Node", id, "(n = "), ""),
        	       info_node(node)$nobs, ifelse(id, ")", ""), sep = "")
      grid.text(mainlab)
      popViewport()

      ## actual plot  
      plot_vpi <- viewport(layout.pos.col = 2, layout.pos.row = 2,
        xscale = xscale, yscale = yscale, 
        name = paste("node_btplot", id, "plot", sep = ""))
      pushViewport(plot_vpi)

      grid.lines(xscale, c(cf_ref, cf_ref), gp = gpar(col = linecol), default.units = "native")
      if(index) {
        grid.lines(x, cfi, gp = gpar(col = col, lty = 2), default.units = "native")
        grid.points(x, cfi, gp = gpar(col = col, cex = cex), pch = pch, default.units = "native")
        grid.xaxis(at = x, label = if(names) names(cfi) else x)
      } else {  	
        if(names) grid.text(names(cfi), x = x, y = cfi, default.units = "native")
          else grid.points(x, cfi, gp = gpar(col = col, cex = cex), pch = pch, default.units = "native")
      }
      grid.yaxis(at = c(ceiling(yscale[1] * 100)/100, floor(yscale[2] * 100)/100))
      grid.rect(gp = gpar(fill = "transparent"))

      upViewport(2)
    }

    return(rval)
}
class(node_btplot) <- "grapcon_generator"
@

\begin{document}

\title{Parties, Models, Mobsters\newline
{\fontsize{11}{15}\selectfont A New Implementation of Model-Based Recursive Partitioning in R}}
\author{Achim Zeileis, Torsten Hothorn}
\URL{http://eeecon.uibk.ac.at/~zeileis/}
\lecture{Parties, Models, Mobsters: A New Implementation of Model-Based Recursive Partitioning in R}

\subsection{Overview}

\begin{frame}
\frametitle{Overview}

\begin{itemize}
  \item Model-based recursive partitioning
  \begin{itemize}
    \item A generic approach
    \item Example: Bradley-Terry trees
  \end{itemize}
  \item Implementation in R
  \begin{itemize}
    \item Building blocks: Parties, models, mobsters
    \item Old implementation in \emph{party}
    \item All new implementation in \emph{partykit}
  \end{itemize}
  \item Illustration
\end{itemize}

\end{frame}

\subsection{Model-based recursive partitioning}

\begin{frame}
\frametitle{Model-based recursive partitioning}

\textbf{Models:} Estimation of parametric models with
observations $y_i$ (and regressors $x_i$),
parameter vector $\theta$, and additive objective function $\Psi$.
\[ \widehat \theta \quad = \quad \argmin_{\theta}
   \sum_i \Psi(y_i, x_i, \theta). \]

\bigskip

\textbf{Recursive partitioning}:

\begin{enumerate}
  \item Fit the model in the current subsample.
  \item Assess the stability of $\theta$ across each partitioning variable $z_j$.
  \item Split sample along the $z_{j^*}$ with strongest association:
        Choose breakpoint with highest improvement of the model fit.
  \item Repeat steps 1--3 recursively in the subsamples
        until some stopping criterion is met.
\end{enumerate}

\end{frame}

\begin{frame}
\frametitle{Model-based recursive partitioning}

\textbf{Parameter instability tests:}
\begin{itemize}
  \item Based on empirical estimating functions (or score/gradient contributions):
    $\Psi'(y_i, x_i, \hat \theta)$.
  \item Under parameter stability: $\Psi'$ fluctuates randomly around its expectation zero.
  \item Under parameter instability: Systematic departures from zero in subsamples.
  \item Hence fluctuation can be captured across numeric partitioning variables
    or within levels of categorical partitioning variables.
  \item Bonferroni correction for testing across multiple partitioning variables.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Bradley-Terry trees}

\hspace*{0.2cm}%
\includegraphics[width=0.23\textwidth]{Barbara.jpg} \hspace*{0.1cm}
\includegraphics[width=0.23\textwidth]{Anni.jpg} \hspace*{0.1cm}
\includegraphics[width=0.23\textwidth]{Hana.jpg} \hspace*{0.1cm}
\includegraphics[width=0.23\textwidth]{Fiona.jpg}

\vspace*{0.1cm}
\hspace*{0.2cm}%
\includegraphics[width=0.23\textwidth]{Mandy.jpg} \hspace*{0.1cm}
\includegraphics[width=0.23\textwidth]{Anja.jpg} \hspace*{0.2cm}%
\begin{tabular}{p{0.48\textwidth}}
\textbf{Questions:} Which of these

women is more attractive?

How does the answer depend on

age, gender, and the familiarity

with the associated TV show

Germany's Next Topmodel?
\vspace*{3.4cm}
\end{tabular}

\end{frame}

\begin{frame}
\frametitle{Bradley-Terry trees}

<<topmodel-tree, echo=FALSE, results=hide>>=
bt0 <- mob(preference ~ 1 | gender + age + q1 + q2 + q3,
  data = Topmodel2007, fit = btfit2)
@

\vspace*{-0.5cm}

\setkeys{Gin}{width=1.15\textwidth}
\hspace*{-0.7cm}%
<<topmodel-plot, echo=FALSE, results=hide, fig=TRUE, height=6.2, width=10>>=
plot(bt0, drop = TRUE, tnex = 2,
  terminal_panel = node_btplot(bt0, abbreviate = 1, yscale = c(0, 0.5)))
@

\end{frame}

\subsection{Implementation in R}

\begin{frame}
\frametitle{Implementation: Building blocks}

\textbf{Workhorse function:} \code{mob()} for
\begin{itemize}
  \item data handling,
  \item calling model fitters,
  \item carrying out parameter instability tests and
  \item recursive partitioning algorithm.
\end{itemize}

\bigskip

\textbf{Required functionality:}
\begin{itemize}
  \item \emph{Parties:} Class and methods for representing recursive partytions.
  \item \emph{Models:} Fitting functions for statistical models (optimizing suitable objective function).
  \item \emph{Mobsters:} High-level interfaces (\code{lmtree()}, \code{bttree()}, \dots)
    that call lower-level \code{mob()} with suitable options and methods.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Implementation: Old \texttt{mob()} in \emph{party}}

\textbf{Parties:} S4 class `\code{BinaryTree}'.
\begin{itemize}
  \item Originally developed only for \code{ctree()} and somewhat ``abused''.
  \item Rather rigid and hard to extend.
\end{itemize}

\medskip

\textbf{Models:} S4 `\code{StatModel}' objects.
\begin{itemize}
  \item Intended to conceptualize unfitted model objects.
  \item Required some ``glue code'' to accomodate non-standard
    interface for data handling and model fitting.
\end{itemize}

\medskip

\textbf{Mobsters:}
\begin{itemize}
  \item \code{mob()} already geared towards (generalized) linear models.
  \item Other interfaces in \emph{psychotree} and \emph{betareg}.
  \item Hard to do fine control due to adopted S4 classes: Many unnecessary
    computations and data storage.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Implementation: New \texttt{mob()} in \emph{partykit}}

\textbf{Parties:} S3 class `\code{modelparty}' built on `\code{party}'.
\begin{itemize}
  \item Separates data and tree structure.
  \item Inherits generic infrastructure for printing, predicting,
    plotting, \dots
\end{itemize}

\medskip

\textbf{Models:} Plain functions with input/output convention.
\begin{itemize}
  \item Basic and extended interface for rapid prototyping and for
    speeding up computings, respectively.
  \item Only minimial glue code required if models are well-designed.
\end{itemize}

\medskip

\textbf{Mobsters:}
\begin{itemize}
  \item \code{mob()} completely agnostic regarding models employed.
  \item Separate interfaces \code{lmtree()}, \code{glmtree()}, \dots
  \item New interfaces typically need to bring their model fitter and adapt the
    main methods \code{print()}, \code{plot()}, \code{predict()} etc.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Implementation: New \texttt{mob()} in \emph{partykit}}

\textbf{New inference options:} Not used by default by optionally available.
\begin{itemize}
  \item New parameter instability tests for ordinal partitioning variables.
    Alternative to unordered $\chi^2$ test but computationally intensive.
  \item Post-pruning based on information criteria (e.g., AIC or BIC), especially
    for very large datasets where traditional significance levels are not useful.
  \item Multiway splits for categorical partitioning variables.
  \item Treat weights as proportionality weights and not as case weights.
\end{itemize}

\end{frame}

\begin{frame}[fragile]
\frametitle{Implementation: Models}

\textbf{Input:} Basic interface.
\begin{Soutput}
  fit(y, x = NULL, start = NULL, weights = NULL,
    offset = NULL, ...)
\end{Soutput}
\code{y}, \code{x}, \code{weights}, \code{offset} are (the subset of)
the preprocessed data.

Starting values and further fitting arguments are in \code{start} and \code{...}.

\bigskip

\textbf{Output:} Fitted model object of class with suitable methods.
\begin{itemize}
  \item \code{coef()}: Estimated parameters $\hat \theta$.
  \item \code{logLik()}: Maximized log-likelihood function $-\sum_i \Psi(y_i, x_, \hat \theta)$.
  \item \code{estfun()}: Empirical estimating functions
    $\Psi'(y_i, x_i, \hat \theta)$.
\end{itemize}

\end{frame}

\begin{frame}[fragile]
\frametitle{Implementation: Models}

\textbf{Input:} Extended interface.
\begin{Soutput}
  fit(y, x = NULL, start = NULL, weights = NULL,
    offset = NULL, ..., estfun = FALSE, object = FALSE)
\end{Soutput}

\medskip

\textbf{Output:} List.
\begin{itemize}
  \item \code{coefficients}: Estimated parameters $\hat \theta$.
  \item \code{objfun}: Minimized objective function $\sum_i \Psi(y_i, x_, \hat \theta)$.
  \item \code{estfun}: Empirical estimating functions
    $\Psi'(y_i, x_i, \hat \theta)$. Only needed if \code{estfun = TRUE}, otherwise optionally \code{NULL}.
  \item \code{object}: A model object for which further methods could
    be available (e.g., \code{predict()}, or \code{fitted()}, etc.).
    Only needed if \code{object = TRUE}, otherwise optionally \code{NULL}.
\end{itemize}

\medskip

\textbf{Internally:} Extended interface constructed from basic interface if supplied.
Efficiency can be gained through extended approach.

\end{frame}

\subsection{Bradley-Terry trees}

\begin{frame}[fragile]
\frametitle{Illustration: Bradley-Terry trees}

Data, packages, and \code{estfun()} method.
<<packages, eval=FALSE>>=
data("Topmodel2007", package = "psychotree")
library("partykit")
library("psychotools")
<<estfun.btReg>>
@

\medskip

Basic model fitting function:

<<btfit1>>=
btfit1 <- function(y, x = NULL, start = NULL, weights = NULL,
  offset = NULL, ...) btReg.fit(y, weights = weights, ...)
@

\medskip

Fit Bradley-Terry tree:

<<bt1>>=
system.time(bt1 <- mob(
  preference ~ 1 | gender + age + q1 + q2 + q3,
  data = Topmodel2007, fit = btfit1))
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Illustration: Bradley-Terry trees}

Extended model fitting function:

<<btfit2-2>>=
<<btfit2>>
@

\medskip

Fit Bradley-Terry tree again:

<<bt2>>=
system.time(bt2 <- mob(
  preference ~ 1 | gender + age + q1 + q2 + q3,
  data = Topmodel2007, fit = btfit2))
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Illustration: Bradley-Terry trees}

<<bt2-print, echo=FALSE>>=
out <- capture.output(bt2)
writeLines(out[1:21])
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Illustration: Bradley-Terry trees}

<<bt2-print, echo=FALSE>>=
out <- capture.output(bt2)
writeLines(out[-(1:22)])
@

\medskip

Standard methods readily available:
<<bt2-coef>>=
plot(bt2)
coef(bt2)
@

\medskip

Customization:
<<bt2-worthf, eval=FALSE>>=
worthf <- function(info) paste(info$object$labels,
  format(round(worth(info$object), digits = 2)), sep = ": ")
plot(bt2, FUN = worthf)
@

\code{}

\end{frame}

\begin{frame}
\frametitle{Illustration: Bradley-Terry trees}

\vspace*{-0.5cm}

\setkeys{Gin}{width=1.15\textwidth}
\hspace*{-0.7cm}%
<<bt2-plot, fig=TRUE, height=7.4, width=12, echo=FALSE>>=
plot(bt2)
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Illustration: Bradley-Terry trees}

\vspace*{-0.5cm}

\setkeys{Gin}{width=1.15\textwidth}
\hspace*{-0.7cm}%
<<bt2-plot2, fig=TRUE, height=7.4, width=12, echo=FALSE>>=
<<bt2-worthf>>
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Illustration: Bradley-Terry trees}

\vspace*{-0.5cm}

\setkeys{Gin}{width=\textwidth}
<<bt2-nodeapply, fig=TRUE, height=7, width=9, echo=FALSE, results=hide>>=
par(mfrow = c(2, 2))
nodeapply(bt2, ids = c(3, 5, 6, 7), FUN = function(n)
  plot(n$info$object, main = n$id, ylim = c(0, 0.4)))
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Illustration: Bradley-Terry trees}

Apply plotting in all terminal nodes:
<<nodeapply, eval=FALSE>>=
<<bt2-nodeapply>>
@

\medskip

<<tm, echo=FALSE>>=
tm <- data.frame(age = c(60, 25, 35), gender = c("male", "female", "female"),
  q1 = "no", q2 = c("no", "no", "yes"), q3 = "no")
@

\medskip

Predicted nodes and ranking:
<<predict>>=
tm
predict(bt2, tm, type = "node")
predict(bt2, tm, type = function(object) t(rank(-worth(object))))
@
%predict(bt2, tm, type = function(object) t(worth(object)))

\end{frame}

\subsection{Summary}

\begin{frame}
\frametitle{Summary}

\begin{itemize}
  \item All new implementation of model-based recursive partitioning in \emph{partykit}.
  \item Enables more efficient computations, rapid prototyping, customization.
  \item Some new inference options.
\end{itemize}

\end{frame}

\subsection{References}

\begin{frame}
\frametitle{References}

\small

Hothorn T, Zeileis A (2014).
 \textit{partykit: A Toolkit for Recursive Partytioning.}
 R~package version~0.2-0.
 URL~\url{http://R-Forge.R-project.org/projects/partykit/}

\bigskip

Zeileis A, Hothorn T (2014).
 \textit{Parties, Models, Mobsters: A New Implementation of Model-Based Recursive Partitioning in R.}
 \code{vignette("mob", package = "partykit")}.

\bigskip

Strobl C, Wickelmaier F, Zeileis A (2011).
  \dquote{Accounting for Individual Differences in Bradley-Terry Models by Means of Recursive Partitioning.}
  \textit{Journal of Educational and Behavioral Statistics}, \textbf{36}(2), 135--153.
  \doi{10.3102/1076998609359791}

\bigskip

Zeileis A, Hothorn T, Hornik K (2008).
 \dquote{Model-Based Recursive Partitioning.}
  \textit{Journal of Computational and Graphical Statistics},
  \textbf{17}(2), 492--514.
  \doi{10.1198/106186008X319331}

\end{frame}

\end{document}