### 2014-02-10 pkg/partykit/inst/NUS-2014/constparty_TH_AZ.Rnw
start constant party talk
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@454 255f4249-2263-4417-b1ea-8498779e27a7

\documentclass[11pt,compress,t]{beamer}

\usetheme{uzhneu-en-informal}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsfonts,amstext,amsmath,amssymb,amsthm}

%% need no 
\usepackage{Sweave}
\definecolor{InputColor}{rgb}{0,0,0.3}
\definecolor{OutputColor}{rgb}{0.2,0.2,0.2}

%% commands
\let\code\texttt
\let\class\texttt
\let\fct\texttt
\let\doi\texttt
\newcommand{\dquote}[1]{``#1''}
\newcommand{\ui}{\underline{i}}
\newcommand{\oi}{\overline{\imath}}
\newcommand{\darrow}{\stackrel{\mbox{\tiny \textnormal{d}}}{\longrightarrow}}
\newcommand{\parrow}{\stackrel{\mbox{\tiny \textnormal{p}}}{\longrightarrow}}
\newcommand{\dotequals}{\stackrel{\cdot}{=}}
\newcommand{\efp}{\mathrm{\it efp}}
\newcommand{\given}{\, | \,}
\newcommand{\ltime}{\lambda_\mathrm{time}}
\newcommand{\lcomp}{\lambda_\mathrm{comp}}
\newcommand{\argmin}{\operatorname{argmin}\displaylimits}

\SweaveOpts{engine=R, eps=FALSE, echo=FALSE, results=hide, keep.source=TRUE}

<<packages>>=
library("partykit")
library("Formula")
library("rpart")
library("RWeka")
library("evtree")
@

<<data>>=
data("Titanic", package = "datasets")
ttnc <- as.data.frame(Titanic)
ttnc <- ttnc[rep(1:nrow(ttnc), ttnc$Freq), 1:4]
names(ttnc)[2] <- "Gender"
ttnc <- transform(ttnc, Treatment = factor(
  Gender == "Female" | Age == "Child",
  levels = c(FALSE, TRUE),
  labels = c("Normal\n(Male&Adult)", "Preferential\n(Female|Child)")
))
@

<<rpart, eval=FALSE>>=
rp <- rpart(Survived ~ Gender + Age + Class, 
            data = ttnc)
@

<<ctree, eval=FALSE>>=
ct <- ctree(Survived ~ Gender + Age + Class, data = ttnc,
  control = ctree_control(mincrit = 1 - 0.01, teststat = "max"))
@

<<evtree, eval=FALSE>>=
set.seed(1071)
ev <- evtree(Survived ~ Gender + Age + Class, 
             data = ttnc, maxdepth = 3)
@

<<pmml, eval=FALSE>>=
pm <- pmmlTreeModel(file.path(system.file("pmml", 
    package = "partykit"), "ttnc.pmml"))
@

<<mob, eval=FALSE>>=
mb <- glmtree(Survived ~ Treatment | Age + Gender + Class,
  data = ttnc, family = binomial, alpha = 0.01)
@

<<cache>>=
if(file.exists("slides-trees.rda")) load("slides-trees.rda") else {
<<rpart>>
<<ctree>>
<<evtree>>
<<pmml>>
<<mob>>
save(rp, ct, ev, pm, mb, file = "slides-trees.rda")
}
@

<<J48>>=
j48 <- J48(Survived ~ Gender + Age + Class, 
           data = ttnc)
@

\begin{document}

\title[Constant Partying]{Constant Partying: \\ 
      Introducing a Generic Toolkit for Recursive Partitioning in R}
\author{Torsten Hothorn (Universit\"at Z\"urich) \\
        Achim Zeileis (Universit\"at Innsbruck)}
%\URL{http://user.math.uzh.ch/hothorn/}
%\lecture{partykit: A Toolkit for Recursive Partytioning}

\maketitle

\begin{frame}
\frametitle{Overview}

\begin{itemize}
  \item Status quo: R software for tree models
  \item New package: partykit
  \begin{itemize}
    \item Unified infrastructure for recursive partytioning
    \item Classes and methods
    \item Interfaces to \code{rpart}, \code{J48}, \dots
    \item Illustrations
  \end{itemize}
  \item Here and now: Intro to partykit for trees with constant fits in each node.
  \item Future: Next steps
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{R software for tree models}

\textbf{Status quo:}
The CRAN task view on ``Machine Learning'' at \url{http://CRAN.R-project.org/view=MachineLearning}
lists numerous packages for tree-based modeling and recursive partitioning, including
%
\begin{itemize}
  \item rpart (CART),
  \item tree (CART),
  \item mvpart (multivariate CART),
  \item RWeka (J4.8, M5', LMT),
  \item party (CTree, MOB),
  \item and many more (C50, quint, stima, \dots).
\end{itemize}

\medskip

\textbf{Related:} Packages for tree-based ensemble methods
such as random forests or boosting, e.g., randomForest, gbm, mboost, etc.

\end{frame}

\begin{frame}
\frametitle{R software for tree models}

\textbf{Furthermore:} Tree algorithms/software without R interface, e.g.,
%
\begin{itemize}
  \item QUEST,
  \item GUIDE,
  \item LOTUS,
  \item CRUISE,
  \item \dots
\end{itemize}

\medskip

\textbf{Currently:}
All algorithms/software have to deal with similar problems \emph{but}
provide different solutions without reusing code.

\end{frame}

\begin{frame}
\frametitle{R software for tree models}

\textbf{Challenge:} For implementing new algorithms in R, code
is required not only for fitting the tree model on the learning data
but also
\begin{itemize}
  \item representing fitted trees,
  \item printing trees,
  \item plotting trees,
  \item computing predictions from trees.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{R software for tree models}

\textbf{Question:} Wouldn't it be nice if there were an R package
that provided code for
\begin{itemize}
  \item representing fitted trees,
  \item printing trees,
  \item plotting trees,
  \item computing predictions from trees?
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{R software for tree models}

\textbf{Answer:} The R package \emph{partykit} provides unified
infrastructure for recursive partytioning, especially
\begin{itemize}
  \item representing fitted trees,
  \item printing trees,
  \item plotting trees,
  \item computing predictions from trees!
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{partykit: Unified infrastructure}

\textbf{Design principles:} Toolkit for recursive partytitioning.
\begin{itemize}
  \item One `agnostic' base class which can encompass an extremely wide
        range of different types of trees.
  \item Subclasses for important types of trees, e.g., trees with constant
        fits in each terminal node.
  \item Nodes are recursive objects, i.e., a node can contain child nodes.
  \item Keep (learning) data out of the recursive node and split structure.
  \item Basic printing, plotting, and predicting for raw node structure.
  \item Customization via suitable panel or panel-generating functions.
  \item Coercion from existing objects (rpart, J48, etc.) to new class.
  \item Use simple/fast S3 classes and methods.
\end{itemize}

\end{frame}

\begin{frame}[fragile]
\frametitle{partykit: Base classes}

\textbf{Class constructors:} Generate basic building blocks.
\begin{itemize}
  \item \code{partysplit(varid, breaks = NULL, index = NULL, ...)}

    where \code{breaks} provides the breakpoints wrt variable \code{varid};
    \code{index} determines to which kid node observations are assigned.

  \item \code{partynode(id, split = NULL, kids = NULL, ...)}

    where \code{split} is a \class{partysplit} and \code{kids} a list of \class{partynode}s.

  \item \code{party(node, data, fitted = NULL, ...)}

    where \code{node} is a \class{partynode} and \code{data} the
    corresponding (learning) data (optionally without any rows) and
    \code{fitted} the corresponding fitted nodes.
\end{itemize}

\medskip

\textbf{Additionally:} All three objects have an \code{info} slot where
optionally arbitrary information can be stored.

\end{frame}

\begin{frame}[fragile]
\frametitle{partykit: Further classes and methods}

\textbf{Further classes:} For trees with constant fits in each terminal node,
both inheriting from \class{party}.
\begin{itemize}
  \item \class{constparty}: Stores full observed response and fitted terminal nodes
    in \code{fitted}; predictions are computed from empirical distribution of
    the response.
  \item \class{simpleparty}: Stores only one predicted response value along with
    some summary details (such as error and sample size) for each terminal node
    in the corresponding \code{info}.
\end{itemize}

\medskip

\textbf{Methods:}
\begin{itemize}
  \item Display: \fct{print}, \fct{plot}, \fct{predict}.
  \item Query: \fct{length}, \fct{width}, \fct{depth}, \fct{names}, \fct{nodeids}.
  \item Extract: \code{[}, \code{[[}, \fct{nodeapply}.
  \item Coercion: \fct{as.party}. 
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{partykit: Illustration}

\textbf{Intention:}
\begin{itemize}
  \item Illustrate several trees using the same data.
  \item Not use the iris data (or something from mlbench).
\end{itemize}

\medskip

\textbf{Solution:}
\begin{itemize}
  \item Use Titanic survival data (oh well\dots).
  \item In case you are not familiar with it: Survival status, gender, age (child/adult), and
    class (1st, 2nd, 3rd, crew) for the \Sexpr{nrow(ttnc)}~persons on the ill-fated
    maiden voyage of the Titanic.
\end{itemize}

\medskip

\textbf{Question:} Who survived? Or how does the probability of survival vary across
the covariates?

\end{frame}

\begin{frame}[fragile]
\frametitle{partykit: Interface to \texttt{rpart}}

\textbf{CART:} Apply \fct{rpart} to preprocessed \code{ttnc} data (see also below).

<<rpart-show, echo=TRUE, eval=FALSE>>=
<<rpart>>
@

\medskip

Standard plot:

<<rpart-plot, echo=TRUE, eval=FALSE>>=
plot(rp)
text(rp)
@

\medskip

Visualization via partykit:

<<rpart-party-plot, echo=TRUE, eval=FALSE>>=
plot(as.party(rp))
@

\end{frame}

\begin{frame}
\frametitle{partykit: Interface to \texttt{rpart}}

\vspace*{-0.5cm}

\setkeys{Gin}{width=1.1\textwidth}
\hspace*{-0.4cm}%
<<Titanic-rpart-visualization1, fig=TRUE, height=6, width=9>>=
<<rpart-plot>>
@

\end{frame}

\begin{frame}
\frametitle{partykit: Interface to \texttt{rpart}}

\vspace*{-0.5cm}

\setkeys{Gin}{width=1.1\textwidth}
\hspace*{-0.4cm}%
<<Titanic-rpart-visualization2, fig=TRUE, height=8, width=12>>=
<<rpart-party-plot>>
@

\end{frame}

\begin{frame}[fragile]
\frametitle{partykit: Interface to \texttt{rpart}}
\small{
<<rpart-print, echo=TRUE, results=verbatim>>=
rp
@
}

\end{frame}

\begin{frame}[fragile]
\frametitle{partykit: Interface to \texttt{rpart}}

\small{
<<rpart-print, echo=TRUE, results=verbatim>>=
as.party(rp)
@
}

\end{frame}

\begin{frame}[fragile]
\frametitle{partykit: Interface to \texttt{rpart}}

\textbf{Prediction:} Compare rpart's C code and partykit's R code

<<rpart-predict, echo=TRUE, results=verbatim>>=
nd <- ttnc[rep(1:nrow(ttnc), 100), ]
system.time(p1 <- predict(rp, newdata = nd, 
                          type = "class"))
system.time(p2 <- predict(as.party(rp), newdata = nd))
table(rpart = p1, party = p2)
@

\end{frame}

\begin{frame}[fragile]
\frametitle{partykit: Interface to \texttt{J48}}

\textbf{J4.8:} Open-source implementation of C4.5 in RWeka.

<<J48-show, echo=TRUE, eval=FALSE>>=
<<J48>>
@

\medskip

Results in a tree with multi-way splits which previously
could only be displayed via Weka itself or Graphviz but not in R
directly. Now:

<<J48-plot, echo=TRUE, eval=FALSE>>=
j48p <- as.party(j48)
plot(j48p)
@

\medskip

Or just a subtree:

<<J48-subtree, echo=TRUE, eval=FALSE>>=
plot(j48p[11])
@

\end{frame}

\begin{frame}
\frametitle{partykit: Interface to \texttt{J48}}

\vspace*{-0.5cm}

\setkeys{Gin}{width=1.1\textwidth}
\hspace*{-0.4cm}%
<<Titanic-J48-visualization, fig=TRUE, height=8, width=12>>=
plot(as.party(j48), tp_args = list(id = FALSE))
@

\end{frame}

\begin{frame}
\frametitle{partykit: Interface to \texttt{J48}}

\vspace*{-0.5cm}

\setkeys{Gin}{width=1.1\textwidth}
\hspace*{-0.4cm}%
<<Titanic-J48-visualization2, fig=TRUE, height=6, width=9>>=
plot(as.party(j48)[11], tnex = 1)
@

\end{frame}

\begin{frame}[fragile]
\frametitle{partykit: Further interfaces}

\textbf{PMML:} Predictive Model Markup Language. XML-based format
exported by various software packages including SAS, SPSS, R/pmml.
Here, import QUEST tree.

<<pmml-show, echo=TRUE, eval=FALSE>>=
<<pmml>>
@

\medskip

\textbf{evtree:} Evolutionary learning of globally optimal trees, directly
using partykit.

<<evtree-show, echo=TRUE, eval=FALSE>>=
<<evtree>>
@

\medskip

\textbf{CTree:} Conditional inference trees \fct{ctree} are reimplemented
more efficiently within partykit.

\medskip

\textbf{CHAID:} R package on R-Forge, directly using partykit.
(Alternatively, use SPSS and export via PMML.)

\end{frame}

\begin{frame}
\frametitle{partykit: QUEST via PMML}

\vspace*{-0.5cm}

\setkeys{Gin}{width=1.1\textwidth}
\hspace*{-0.4cm}%
<<Titanic-pmml-visualization, fig=TRUE, height=8, width=12>>=
plot(pm)
@

\end{frame}

\begin{frame}
\frametitle{partykit: evtree}

\vspace*{-0.5cm}

\setkeys{Gin}{width=1.1\textwidth}
\hspace*{-0.4cm}%
<<Titanic-evtree-visualization, fig=TRUE, height=8, width=12>>=
plot(ev)
@

\end{frame}

\begin{frame}
\frametitle{Your own tree}

\end{frame}

\begin{frame}
\frametitle{What's next?}

\textbf{Model-based recursive partitioning:} Trees with parametric models in each node
(e.g., based on least squares or maximum likelihood). Splitting based on parameter
instability tests.

\medskip

Talk by Achim; but I'm entitled to a coffee first...

\end{frame}

\begin{frame}[fragile]
\frametitle{Computational details}

\textbf{Software:} All examples have been produced with
R~\Sexpr{with(R.Version(), paste(major, minor, sep = "."))} and packages
partykit~\Sexpr{packageDescription("partykit")$Version},
rpart~\Sexpr{packageDescription("rpart")$Version},
RWeka~\Sexpr{packageDescription("RWeka")$Version},
evtree~\Sexpr{packageDescription("evtree")$Version}. All packages are
freely available under the GPL from \url{http://CRAN.R-project.org/}.

\bigskip

\textbf{Data:} The contingency table \code{Titanic} is transformed
as explained in the \code{constparty} package vignette.

\end{frame}

\begin{frame}
\frametitle{References}

\small

Hothorn T, Zeileis A (2013).
 \textit{partykit: A Toolkit for Recursive Partytioning.}
 R~package vignette version~0.1-6.
 URL~\url{http://CRAN.R-project.org/package=partykit}

\bigskip

Hothorn T, Hornik K, Zeileis A (2006).
 \dquote{Unbiased Recursive Partitioning: A Conditional Inference Framework.}
 \textit{Journal of Computational and Graphical Statistics},
 \textbf{15}(3), 651--674.
 \doi{10.1198/106186006X133933}

\bigskip

Grubinger T, Zeileis A, Pfeiffer KP (2011).
 \dquote{evtree: Evolutionary Learning of Globally Optimal Classification and Regression Trees in R.}
 \textit{Working Paper~2011-20}, Working Papers in Economics and Statistics,
 Research Platform Empirical and Experimental Economics, Universit\"at Innsbruck.
 URL~\url{http://EconPapers.RePEc.org/RePEc:inn:wpaper:2011-20}

\bigskip

Zeileis A, Hothorn T, Hornik K (2008).
 \dquote{Model-Based Recursive Partitioning.}
  \textit{Journal of Computational and Graphical Statistics},
  \textbf{17}(2), 492--514.
  \doi{10.1198/106186008X319331}

\end{frame}

\end{document}