### 2014-03-09 pkg/partykit/inst/NUS-2014/modelparty.Rnw
merged slides from JKU and Psychoco -> still needs a lot of work
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@465 255f4249-2263-4417-b1ea-8498779e27a7

\documentclass[11pt,compress,t]{beamer}
\usetheme{Z}
\usepackage{amsfonts,amstext,amsmath}
%% need no \usepackage{Sweave}
\definecolor{InputColor}{rgb}{0,0,0.3}
\definecolor{OutputColor}{rgb}{0.2,0.2,0.2}

\graphicspath{{../Psychoco-2014/images/}}

%% commands
\newcommand{\ui}{\underline{i}}
\newcommand{\oi}{\overline{\imath}}
\newcommand{\darrow}{\stackrel{\mbox{\tiny \textnormal{d}}}{\longrightarrow}}
\newcommand{\parrow}{\stackrel{\mbox{\tiny \textnormal{p}}}{\longrightarrow}}
\newcommand{\dotequals}{\stackrel{\cdot}{=}}
\newcommand{\efp}{\mathrm{\it efp}}
\newcommand{\given}{\, | \,}
\newcommand{\ltime}{\lambda_\mathrm{time}}
\newcommand{\lcomp}{\lambda_\mathrm{comp}}
\newcommand{\argmin}{\operatorname{argmin}\displaylimits}

\SweaveOpts{engine=R, eps=FALSE, keep.source=TRUE}

<<packages, echo=FALSE, results=hide>>=
library("partykit")
library("psychotools")
data("Topmodel2007", package = "psychotree")
options(digits = 4)
@

<<btfit2, echo=FALSE, results=hide>>=
btfit2 <- function(y, x = NULL, start = NULL, weights = NULL,
  offset = NULL, ..., estfun = FALSE, object = FALSE) {
  rval <- btReg.fit(y, weights = weights, ...,
    estfun = estfun, vcov = object)
  list(
    coefficients = rval$coefficients,
    objfun = -rval$loglik,
    estfun = if(estfun) rval$estfun else NULL,
    object = if(object) rval else NULL
  )
}
@

<<estfun.btReg, echo=FALSE, results=hide>>=
estfun.btReg <- function(x, ...) x$estfun
@

<<panelfunction, echo=FALSE, results=hide>>=
## visualization function
node_btplot <- function(mobobj, id = TRUE,
  worth = TRUE, names = TRUE, abbreviate = TRUE, index = TRUE, ref = TRUE,
  col = "black", linecol = "lightgray", cex = 0.5, pch = 19, xscale = NULL, yscale = NULL, ylines = 1.5)
{
    ## node ids
    node <- nodeids(mobobj, terminal = FALSE)

    ## get all coefficients 
    cf <- partykit:::apply_to_models(mobobj, node, FUN = function(z)        
      if(worth) worth(z) else coef(z, all = FALSE, ref = TRUE))
    cf <- do.call("rbind", cf)
    rownames(cf) <- node

    ## get one full model
    mod <- partykit:::apply_to_models(mobobj, node = 1L, FUN = NULL)

    if(!worth) {
      if(is.character(ref) | is.numeric(ref)) {
        reflab <- ref
        ref <- TRUE
      } else {
        reflab <- mod$ref
      }
      if(is.character(reflab)) reflab <- match(reflab, mod$labels)
      cf <- cf - cf[,reflab]
    }

    ## reference
    if(worth) {
      cf_ref <- 1/ncol(cf)
    } else {
      cf_ref <- 0
    }

    ## labeling
    if(is.character(names)) {
      colnames(cf) <- names
      names <- TRUE
    }

    ## abbreviation
    if(is.logical(abbreviate)) {
      nlab <- max(nchar(colnames(cf)))
      abbreviate <- if(abbreviate) as.numeric(cut(nlab, c(-Inf, 1.5, 4.5, 7.5, Inf))) else nlab
    }
    colnames(cf) <- abbreviate(colnames(cf), abbreviate)

    if(index) {
      x <- 1:NCOL(cf)
      if(is.null(xscale)) xscale <- range(x) + c(-0.1, 0.1) * diff(range(x))
    } else {
      x <- rep(0, length(cf))
      if(is.null(xscale)) xscale <- c(-1, 1)      
    }
    if(is.null(yscale)) yscale <- range(cf) + c(-0.1, 0.1) * diff(range(cf))

    ## panel function for bt plots in nodes
    rval <- function(node) {

      ## node index
      id <- id_node(node)

      ## dependent variable setup
      cfi <- cf[id,]

      ## viewport setup
      top_vp <- viewport(layout = grid.layout(nrow = 2, ncol = 3,
    			 widths = unit(c(ylines, 1, 1), c("lines", "null", "lines")),  
        		 heights = unit(c(1, 1), c("lines", "null"))),
    			 width = unit(1, "npc"), 
    			 height = unit(1, "npc") - unit(2, "lines"),
        		 name = paste("node_btplot", id, sep = ""))
      pushViewport(top_vp)
      grid.rect(gp = gpar(fill = "white", col = 0))

      ## main title
      top <- viewport(layout.pos.col = 2, layout.pos.row = 1)
      pushViewport(top)
      mainlab <- paste(ifelse(id, paste("Node", id, "(n = "), ""),
        	       info_node(node)$nobs, ifelse(id, ")", ""), sep = "")
      grid.text(mainlab)
      popViewport()

      ## actual plot  
      plot_vpi <- viewport(layout.pos.col = 2, layout.pos.row = 2,
        xscale = xscale, yscale = yscale, 
        name = paste("node_btplot", id, "plot", sep = ""))
      pushViewport(plot_vpi)

      grid.lines(xscale, c(cf_ref, cf_ref), gp = gpar(col = linecol), default.units = "native")
      if(index) {
        grid.lines(x, cfi, gp = gpar(col = col, lty = 2), default.units = "native")
        grid.points(x, cfi, gp = gpar(col = col, cex = cex), pch = pch, default.units = "native")
        grid.xaxis(at = x, label = if(names) names(cfi) else x)
      } else {  	
        if(names) grid.text(names(cfi), x = x, y = cfi, default.units = "native")
          else grid.points(x, cfi, gp = gpar(col = col, cex = cex), pch = pch, default.units = "native")
      }
      grid.yaxis(at = c(ceiling(yscale[1] * 100)/100, floor(yscale[2] * 100)/100))
      grid.rect(gp = gpar(fill = "transparent"))

      upViewport(2)
    }

    return(rval)
}
class(node_btplot) <- "grapcon_generator"
@

\begin{document}

\title{Parties, Models, Mobsters\newline
{\fontsize{11}{15}\selectfont A New Implementation of Model-Based Recursive Partitioning in R}}
\author{Achim Zeileis, Torsten Hothorn}
\URL{http://eeecon.uibk.ac.at/~zeileis/}
\lecture{Parties, Models, Mobsters: A New Implementation of Model-Based Recursive Partitioning in R}

\subsection{Overview}

\begin{frame}
\frametitle{Overview}

\begin{itemize}
  \item Model-based recursive partitioning
  \begin{itemize}
    \item A generic approach
    \item Example: Bradley-Terry trees
  \end{itemize}
  \item Implementation in R
  \begin{itemize}
    \item Building blocks: Parties, models, mobsters
    \item Old implementation in \emph{party}
    \item All new implementation in \emph{partykit}
  \end{itemize}
  \item Illustration
\end{itemize}

\end{frame}

\subsection{Motivation}

\begin{frame}
\frametitle{Motivation: Trees}

Breiman (2001, \textit{Statistical Science}) distinguishes two cultures
of statistical modeling.

\medskip

\begin{itemize}
  \item \textbf{Data models:} Stochastic models, typically parametric.\\
        $\rightarrow$ Classical strategy in statistics.
	Regression models are still the workhorse for many empirical analyses.
  \item \textbf{Algorithmic models:} Flexible models, data-generating process unknown.
        $\rightarrow$ Still few applications in many fields, e.g., social sciences or economics.
\end{itemize}

\medskip

\textbf{Examples:} Recursive partitioning, decision trees.
Class of flexible methods for classification and regression.

\end{frame}

\begin{frame}
\frametitle{Motivation: Trees}

\textbf{Formally}: Modeling of dependent variable $Y$ by ``learning''
a recursive partition w.r.t\ explanatory variables $Z_1, \dots, Z_l$.

\bigskip

\textbf{Key features}:

\begin{enumerate}
  \item Predictive power in nonlinear regression relationships.
  \item Interpretability (enhanced by visualization), i.e., no ``black box'' methods.
\end{enumerate}

\bigskip

\textbf{Examples}: CART and C4.5 in statistical and machine learning, respectively.

\end{frame}

\begin{frame}
\frametitle{Motivation: Leaves}

\textbf{Typically:} Simple models for univeriate $Y$, e.g., mean.

\bigskip

\textbf{Idea:} More complex models for multivariate $Y$, e.g., multivariate
normal model, regression models, etc.

\bigskip

\textbf{Here:} Synthesis of parametric data models and algorithmic tree models.

\bigskip

\textbf{Goal:} Fitting local models by partitioning of the sample space.

\end{frame}

\subsection{Recursive partitioning}

\begin{frame}
\frametitle{Recursive partitioning}

\textbf{Base algorithm}:

\begin{enumerate}
  \item Fit model for $Y$.
  \item Assess association of $Y$ and each $Z_j$.
  \item Split sample along the $Z_{j^*}$ with strongest association:
        Choose breakpoint with highest improvement of the model fit.
  \item Repeat steps 1--3 recursively in the subsamples
        until some stopping criterion is met.
\end{enumerate}

%% \medskip
%% 
%% \textbf{Generally:} Tree algorithms differ w.r.t.\ choice of model (1),
%% association measure~(2), split strategy (3) and stopping criterion (4).

\bigskip

\textbf{Here:} Segmentation~(3) of parametric models~(1) with 
additive objective function using parameter instability tests~(2)
and associated statistical significance~(4).

\end{frame}

\subsection{Model estimation}

\begin{frame}
\frametitle{1.~Model estimation}

\textbf{Models:} $\mathcal{M}(Y, \theta)$ with (potentially) multivariate
observations $Y \in \mathcal{Y}$ and $k$-dimensional parameter vector
$\theta \in \Theta$.

\bigskip

\textbf{Parameter estimation:} $\widehat \theta$ by optimization of
objective function $\Psi(Y, \theta)$ for $n$ observations $Y_i$ ($i = 1, \dots, n$):

\[ \widehat \theta \quad = \quad \argmin_{\theta \in \Theta}
   \sum_{i = 1}^n \Psi(Y_i, \theta). \]

\bigskip

\textbf{Special cases:} Maximum likelihood (ML), weighted and ordinary
least squares (OLS and WLS), quasi-ML, and other M-estimators.

\end{frame}

\begin{frame}
\frametitle{1.~Model estimation}

\textbf{Estimating function:} $\widehat{\theta}$ can also be defined in terms of
  \[ \sum_{i = 1}^n \psi(Y_i, \widehat{\theta}) = 0, \]
where $\psi(Y, \theta) = {\partial \Psi(Y, \theta)}/{\partial \theta}$.

\bigskip

\textbf{Central limit theorem:} If there is a true parameter $\theta_0$ and given
certain weak regularity conditions:
  \[ \sqrt{n} (\widehat{\theta} - \theta_0) \quad \darrow \quad \mathcal{N}(0, V(\theta_0)), \]
where $V(\theta_0) = \{A(\theta_0)\}^{-1} B(\theta_0) \{A(\theta_0)\}^{-1}$.
$A$ and $B$ are the expectation of the derivative of $\psi$ and
the variance of $\psi$, respectively.

\end{frame}

\begin{frame}
\frametitle{1.~Model estimation}

\textbf{Idea:}
In many situations, a single global model 
$\mathcal{M}(Y, \theta)$ that fits \textbf{all} $n$~observations cannot be found.
But it might be possible to find a partition w.r.t.\ the variables
$Z = (Z_1, \dots, Z_l)$ so that a well-fitting model can be found locally
in each cell of the partition.

\bigskip

\textbf{Tool:} Assess parameter instability w.r.t\ to partitioning variables
$Z_j \in \mathcal{Z}_j \; (j = 1, \dots, l)$.

\end{frame}

\subsection{Parameter instability tests}

\begin{frame}
\frametitle{2.~Tests for parameter instability}

Generalized M-fluctuation tests
capture instabilities in $\widehat \theta$ for an ordering w.r.t\ $Z_j$.

\bigskip

\textbf{Basis:} Empirical fluctuation process of cumulative deviations
w.r.t.\ to an ordering $\sigma(Z_{ij})$.

\[
W_j(t, \widehat \theta)
  \quad = \quad {\widehat B}^{-1/2} n^{-1/2}
                \sum_{i = 1}^{\lfloor nt \rfloor}
		\psi(Y_{\sigma(Z_{ij})}, \widehat \theta)
\qquad (0 \le t \le 1)
\]

\bigskip

\textbf{Functional central limit theorem:} Under parameter stability
$W_j(\cdot) \, \darrow \, W^0(\cdot)$, where
$W^0$ is a $k$-dimensional Brownian bridge.

\end{frame}

\begin{frame}
\frametitle{2.~Tests for parameter instability}

\setkeys{Gin}{width=1.08\textwidth}
\hspace*{-0.7cm}%
<<fig=TRUE, height=5, width=10, echo=FALSE, results=hide>>=
## artificial data
set.seed(42)
a <- rnorm(41, mean = 0, sd = 0.12)
b <- rnorm(41, mean = 1, sd = 0.12)
b[40] <- b[40]-0.1
b[41] <- b[41]-0.2
a[1] <- a[1]+0.2
a[2] <- a[2]+0.1
x <- c(b, a)

par(mar = c(5, 5, 3, 1), mfrow = c(1, 2))

## data, model, and residuals
plot(0:81, x, xlab = expression(Z[j]), ylab = "Y", ylim = c(-2, 3),
  type = "l", xlim = c(1, 80), axes=FALSE)
axis(1,at=seq(0,80,20),labels=seq(2004,2012,2),cex.axis=1.2, cex.lab=1.2)
axis(2,at=seq(-2,3,1),labels=seq(200,1200,200),cex.axis=1.2, cex.lab=1.2)
box()
abline(h = mean(x), lty = 3)
for(i in c(5, 15, 25, 35)) {
  segments(x0 = i, x1 = i, y0 = x[i+1], y1 = mean(x), lty = 2)
  i <- i + 40
  segments(x0 = i, x1 = i, y0 = mean(x), y1 = x[i+1], lty = 2)
}

## CUSUM graph
x_efp <- c(0, cumsum(x - mean(x)))/sqrt(length(x) * var(x))
plot(-1:81, x_efp, xlab = expression(Z[j]), ylab = "Fluctuation Process", 
  type = "l", xlim = c(1, 80), ylim = c(-5, 5), axes=FALSE)
axis(1,at=seq(0,80,20),labels=seq(2004,2012,2),cex.axis=1.2, cex.lab=1.2)
axis(2, cex.axis=1.2, cex.lab=1.2)
box()
abline(h = 0, lty = 3)
for(i in seq(5, 75, by = 10)) segments(x0 = i, x1 = i, y0 = 0, y1 = x_efp[i+2], lty = 2)
@

\end{frame}

\begin{frame}
\frametitle{2.~Tests for parameter instability}

\textbf{Test statistics:} Scalar functional $\lambda(W_j)$
that captures deviations from zero.

\bigskip

\textbf{Null distribution:} Asymptotic distribution of $\lambda(W^0)$.

\bigskip

\textbf{Special cases:} Class of test encompasses many well-known
tests for different classes of models.
Certain functionals~$\lambda$ are particularly intuitive for numeric and
categorical $Z_j$, respectively.

\bigskip

\textbf{Advantage:} Model $\mathcal{M}(Y, \widehat \theta)$ just has to be
estimated once. Empirical estimating functions $\psi(Y_i, \widehat \theta)$
just have to be re-ordered and aggregated for each $Z_j$.

\end{frame}

\begin{frame}
\frametitle{2.~Tests for parameter instability}

\textbf{Splitting numeric variables:} Assess instability using
sup$\mathit{LM}$ statistics.

\[
\lambda_{\mbox{sup}\mathit{LM}} (W_j) \quad = \quad
\max_{i = \ui, \dots, \oi}  \, \left(\frac{i}{n} \cdot \frac{n-i}{n} \right)^{-1}
\left| \left| W_j \left( \frac{i}{n} \right) \right| \right|^2_2.
\]

\bigskip

\textbf{Interpretation:} Maximization of single shift $\mathit{LM}$ statistics
for all conceivable breakpoints in $[\ui, \oi]$.

\bigskip

\textbf{Limiting distribution:} Supremum of a squared, $k$-dimensional tied-down
Bessel process.

\end{frame}

\begin{frame}
\frametitle{2.~Tests for parameter instability}

\textbf{Splitting categorical variables:} Assess instability using
$\chi^2$ statistics.

\[ \lambda_{\chi^2} (W_j) \quad = \quad
\sum_{c = 1}^C 
\frac{n}{|I_c|}  \left| \left| \Delta_{I_c} W_j \left( \frac{i}{n} \right) \right| \right|^2_2
\]

\bigskip

\textbf{Feature:} Invariant for re-ordering of the $C$ categories and the
observations within each category.

\bigskip

\textbf{Interpretation:} Captures instability for split-up into $C$ categories.

\bigskip

\textbf{Limiting distribution:} $\chi^2$ with $k \cdot (C-1)$ degrees of freedom.

\end{frame}

\begin{frame}
\frametitle{2.~Tests for parameter instability}

\textbf{Splitting ordinal variables:} FIXME

like categorical

like numeric but invariant to reordering within ties

\end{frame}

\subsection{Segmentation}

\begin{frame}
\frametitle{3.~Segmentation}

\textbf{Goal:} Split model into $b = 1, \dots, B$ segments along the
partitioning variable $Z_j$ associated with the highest parameter instability.
Local optimization of

\[ \sum_b \sum_{i \in I_b} \Psi(Y_i, \theta_b). \]

\bigskip

$B = 2$: Exhaustive search of order $O(n)$.

\bigskip

$B > 2$: Exhaustive search is of order $O(n^{B-1})$, but can be replaced
by dynamic programming of order $O(n^2)$. Different methods (e.g., information
criteria) can choose $B$ adaptively.

\bigskip

\textbf{Here:} Binary partitioning.

\end{frame}

\subsection{Pruning}

\begin{frame}
\frametitle{4.~Pruning}

\textbf{Pruning:} Avoid overfitting.

\bigskip

\textbf{Pre-pruning:} Internal stopping criterium. Stop splitting when
there is no significant parameter instability based on
Bonferroni-corrected $p$~values of the fluctuation tests.

\bigskip

\textbf{Post-pruning:} Grow large tree and prune splits that do not
improve the model fit based on information criteria (e.g., AIC or BIC).

\end{frame}

\begin{frame}
\frametitle{Bradley-Terry trees}

\hspace*{0.2cm}%
\includegraphics[width=0.23\textwidth]{Barbara.jpg} \hspace*{0.1cm}
\includegraphics[width=0.23\textwidth]{Anni.jpg} \hspace*{0.1cm}
\includegraphics[width=0.23\textwidth]{Hana.jpg} \hspace*{0.1cm}
\includegraphics[width=0.23\textwidth]{Fiona.jpg}

\vspace*{0.1cm}
\hspace*{0.2cm}%
\includegraphics[width=0.23\textwidth]{Mandy.jpg} \hspace*{0.1cm}
\includegraphics[width=0.23\textwidth]{Anja.jpg} \hspace*{0.2cm}%
\begin{tabular}{p{0.48\textwidth}}
\textbf{Questions:} Which of these

women is more attractive?

How does the answer depend on

age, gender, and the familiarity

with the associated TV show

Germany's Next Topmodel?
\vspace*{3.4cm}
\end{tabular}

\end{frame}

\begin{frame}
\frametitle{Bradley-Terry trees}

<<topmodel-tree, echo=FALSE, results=hide>>=
bt0 <- mob(preference ~ 1 | gender + age + q1 + q2 + q3,
  data = Topmodel2007, fit = btfit2)
@

\vspace*{-0.5cm}

\setkeys{Gin}{width=1.15\textwidth}
\hspace*{-0.7cm}%
<<topmodel-plot, echo=FALSE, results=hide, fig=TRUE, height=6.2, width=10>>=
plot(bt0, drop = TRUE, tnex = 2,
  terminal_panel = node_btplot(bt0, abbreviate = 1, yscale = c(0, 0.5)))
@

\end{frame}

\subsection{Implementation in R}

\begin{frame}
\frametitle{Implementation: Building blocks}

\textbf{Workhorse function:} \code{mob()} for
\begin{itemize}
  \item data handling,
  \item calling model fitters,
  \item carrying out parameter instability tests and
  \item recursive partitioning algorithm.
\end{itemize}

\bigskip

\textbf{Required functionality:}
\begin{itemize}
  \item \emph{Parties:} Class and methods for recursive partytions.
  \item \emph{Models:} Fitting functions for statistical models (optimizing suitable objective function).
  \item \emph{Mobsters:} High-level interfaces (\code{lmtree()}, \code{bttree()}, \dots)
    that call lower-level \code{mob()} with suitable options and methods.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Implementation: Old \texttt{mob()} in \emph{party}}

\textbf{Parties:} S4 class `\code{BinaryTree}'.
\begin{itemize}
  \item Originally developed only for \code{ctree()} and somewhat ``abused''.
  \item Rather rigid and hard to extend.
\end{itemize}

\medskip

\textbf{Models:} S4 `\code{StatModel}' objects.
\begin{itemize}
  \item Intended to conceptualize unfitted model objects.
  \item Required some ``glue code'' to accomodate non-standard
    interface for data handling and model fitting.
\end{itemize}

\medskip

\textbf{Mobsters:}
\begin{itemize}
  \item \code{mob()} already geared towards (generalized) linear models.
  \item Other interfaces in \emph{psychotree} and \emph{betareg}.
  \item Hard to do fine control due to adopted S4 classes: Many unnecessary
    computations and copies of data.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Implementation: New \texttt{mob()} in \emph{partykit}}

\textbf{Parties:} S3 class `\code{modelparty}' built on `\code{party}'.
\begin{itemize}
  \item Separates data and tree structure.
  \item Inherits generic infrastructure for printing, predicting,
    plotting, \dots
\end{itemize}

\medskip

\textbf{Models:} Plain functions with input/output convention.
\begin{itemize}
  \item Basic and extended interface for rapid prototyping and for
    speeding up computings, respectively.
  \item Only minimial glue code required if models are well-designed.
\end{itemize}

\medskip

\textbf{Mobsters:}
\begin{itemize}
  \item \code{mob()} completely agnostic regarding models employed.
  \item Separate interfaces \code{lmtree()}, \code{glmtree()}, \dots
  \item New interfaces typically need to bring their model fitter and adapt the
    main methods \code{print()}, \code{plot()}, \code{predict()} etc.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Implementation: New \texttt{mob()} in \emph{partykit}}

\textbf{New inference options:} Not used by default by optionally available.
\begin{itemize}
  \item New parameter instability tests for ordinal partitioning variables.
    Alternative to unordered $\chi^2$ test but computationally intensive.
  \item Post-pruning based on information criteria (e.g., AIC or BIC), especially
    for very large datasets where traditional significance levels are not useful.
  \item Multiway splits for categorical partitioning variables.
  \item Treat weights as proportionality weights and not as case weights.
\end{itemize}

\end{frame}

\begin{frame}[fragile]
\frametitle{Implementation: Models}

\textbf{Input:} Basic interface.
\begin{Soutput}
  fit(y, x = NULL, start = NULL, weights = NULL,
    offset = NULL, ...)
\end{Soutput}
\code{y}, \code{x}, \code{weights}, \code{offset} are (the subset of)
the preprocessed data.

Starting values and further fitting arguments are in \code{start} and \code{...}.

\bigskip

\textbf{Output:} Fitted model object of class with suitable methods.
\begin{itemize}
  \item \code{coef()}: Estimated parameters $\hat \theta$.
  \item \code{logLik()}: Maximized log-likelihood function $-\sum_i \Psi(y_i, x_, \hat \theta)$.
  \item \code{estfun()}: Empirical estimating functions
    $\Psi'(y_i, x_i, \hat \theta)$.
\end{itemize}

\end{frame}

\begin{frame}[fragile]
\frametitle{Implementation: Models}

\textbf{Input:} Extended interface.
\begin{Soutput}
  fit(y, x = NULL, start = NULL, weights = NULL,
    offset = NULL, ..., estfun = FALSE, object = FALSE)
\end{Soutput}

\medskip

\textbf{Output:} List.
\begin{itemize}
  \item \code{coefficients}: Estimated parameters $\hat \theta$.
  \item \code{objfun}: Minimized objective function $\sum_i \Psi(y_i, x_, \hat \theta)$.
  \item \code{estfun}: Empirical estimating functions
    $\Psi'(y_i, x_i, \hat \theta)$. Only needed if \code{estfun = TRUE}, otherwise optionally \code{NULL}.
  \item \code{object}: A model object for which further methods could
    be available (e.g., \code{predict()}, or \code{fitted()}, etc.).
    Only needed if \code{object = TRUE}, otherwise optionally \code{NULL}.
\end{itemize}

\medskip

\textbf{Internally:} Extended interface constructed from basic interface if supplied.
Efficiency can be gained through extended approach.

\end{frame}

\subsection{Bradley-Terry trees}

\begin{frame}[fragile]
\frametitle{Illustration: Bradley-Terry trees}

Data, packages, and \code{estfun()} method:
<<packages, eval=FALSE>>=
data("Topmodel2007", package = "psychotree")
library("partykit")
library("psychotools")
<<estfun.btReg>>
@

\medskip

Basic model fitting function:

<<btfit1>>=
btfit1 <- function(y, x = NULL, start = NULL, weights = NULL,
  offset = NULL, ...) btReg.fit(y, weights = weights, ...)
@

\medskip

Fit Bradley-Terry tree:

<<bt1>>=
system.time(bt1 <- mob(
  preference ~ 1 | gender + age + q1 + q2 + q3,
  data = Topmodel2007, fit = btfit1))
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Illustration: Bradley-Terry trees}

Extended model fitting function:

<<btfit2-2>>=
<<btfit2>>
@

\medskip

Fit Bradley-Terry tree again:

<<bt2>>=
system.time(bt2 <- mob(
  preference ~ 1 | gender + age + q1 + q2 + q3,
  data = Topmodel2007, fit = btfit2))
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Illustration: Bradley-Terry trees}

<<bt2-print, echo=FALSE>>=
out <- capture.output(bt2)
writeLines(out[1:21])
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Illustration: Bradley-Terry trees}

<<bt2-print, echo=FALSE>>=
out <- capture.output(bt2)
writeLines(out[-(1:22)])
@

\medskip

Standard methods readily available:
<<bt2-coef>>=
plot(bt2)
coef(bt2)
@

\medskip

Customization:
<<bt2-worthf, eval=FALSE>>=
worthf <- function(info) paste(info$object$labels,
  format(round(worth(info$object), digits = 2)), sep = ": ")
plot(bt2, FUN = worthf)
@

\code{}

\end{frame}

\begin{frame}
\frametitle{Illustration: Bradley-Terry trees}

\vspace*{-0.5cm}

\setkeys{Gin}{width=1.15\textwidth}
\hspace*{-0.7cm}%
<<bt2-plot, fig=TRUE, height=7.4, width=12, echo=FALSE>>=
plot(bt2)
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Illustration: Bradley-Terry trees}

\vspace*{-0.5cm}

\setkeys{Gin}{width=1.15\textwidth}
\hspace*{-0.7cm}%
<<bt2-plot2, fig=TRUE, height=7.4, width=12, echo=FALSE>>=
<<bt2-worthf>>
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Illustration: Bradley-Terry trees}

\vspace*{-0.5cm}

\setkeys{Gin}{width=\textwidth}
<<bt2-nodeapply, fig=TRUE, height=7, width=9, echo=FALSE, results=hide>>=
par(mfrow = c(2, 2))
nodeapply(bt2, ids = c(3, 5, 6, 7), FUN = function(n)
  plot(n$info$object, main = n$id, ylim = c(0, 0.4)))
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Illustration: Bradley-Terry trees}

Apply plotting in all terminal nodes:
<<nodeapply, eval=FALSE>>=
<<bt2-nodeapply>>
@

\medskip

<<tm, echo=FALSE>>=
tm <- data.frame(age = c(60, 25, 35), gender = c("male", "female", "female"),
  q1 = "no", q2 = c("no", "no", "yes"), q3 = "no")
@

\medskip

Predicted nodes and ranking:
<<predict>>=
tm
predict(bt2, tm, type = "node")
predict(bt2, tm, type = function(object) t(rank(-worth(object))))
@
%predict(bt2, tm, type = function(object) t(worth(object)))

\end{frame}

\subsection{Summary}

\begin{frame}
\frametitle{Summary}

\begin{itemize}
  \item All new implementation of model-based recursive partitioning in \emph{partykit}.
  \item Enables more efficient computations, rapid prototyping, flexible customization.
  \item Some new inference options.
\end{itemize}

\end{frame}

\subsection{References}

\begin{frame}
\frametitle{References}

\small

Hothorn T, Zeileis A (2014).
 \textit{partykit: A Toolkit for Recursive Partytioning.}
 R~package version~0.2-0.
 URL~\url{http://R-Forge.R-project.org/projects/partykit/}

\bigskip

Zeileis A, Hothorn T (2014).
 \textit{Parties, Models, Mobsters: A New Implementation of Model-Based Recursive Partitioning in R.}
 \code{vignette("mob", package = "partykit")}.

\bigskip

Strobl C, Wickelmaier F, Zeileis A (2011).
  \dquote{Accounting for Individual Differences in Bradley-Terry Models by Means of Recursive Partitioning.}
  \textit{Journal of Educational and Behavioral Statistics}, \textbf{36}(2), 135--153.
  \doi{10.3102/1076998609359791}

\bigskip

Zeileis A, Hothorn T, Hornik K (2008).
 \dquote{Model-Based Recursive Partitioning.}
  \textit{Journal of Computational and Graphical Statistics},
  \textbf{17}(2), 492--514.
  \doi{10.1198/106186008X319331}

\end{frame}

\end{document}