### 2014-03-19 pkg/partykit/inst/mloss/mloss2.Rnw
start assembling 4-page paper for MLOSS 2014
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@509 255f4249-2263-4417-b1ea-8498779e27a7

\documentclass[a4paper]{article}
\usepackage{jmlr2e}
\usepackage{color,hyperref}

%% software markup
\let\proglang=\textsf
\let\pkg=\textit
\let\code=\texttt
\newcommand{\squote}[1]{`{#1}'}
\newcommand{\dquote}[1]{``{#1}''}
\newcommand{\fct}[1]{\texttt{#1()}}
\newcommand{\class}[1]{\squote{\texttt{#1}}}

%% \usepackage{Sweave} is essentially
\RequirePackage[T1]{fontenc}
\RequirePackage{ae,fancyvrb}
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
\newenvironment{Schunk}{}{}
\setkeys{Gin}{width=.7\textwidth}
\SweaveOpts{engine=R, eps=FALSE, keep.source=TRUE, echo=FALSE, results=hide}
<<options, echo=FALSE>>=
options(prompt = "R> ", continue = "+  ", digits = 4, width = 70)
library("partykit")
set.seed(290875)
@

%% header declarations
\author{Torsten Hothorn \\ \addr{Universit\"at Z\"urich} \\ \addr{E-mail: Torsten.Hothorn@R-project.org} \AND
        Achim Zeileis \\ \addr{Universit\"at Innsbruck} \\ \addr{E-mail: Achim.Zeileis@R-project.org}}
\title{\pkg{partykit}: An Open-Source Toolkit for Recursive Partytioning in \proglang{R}}
\editor{(submission for MLOSS track)}

% hyperref setup
\definecolor{Red}{rgb}{0.5,0,0}
\definecolor{Blue}{rgb}{0,0,0.5}
\hypersetup{%
  pdftitle = {partykit: An Open-Source Toolkit for Recursive Partytioning in R},
  pdfsubject = {submitted to JMLR-MLOSS},
  pdfkeywords = {open-source software, recursive partitioning, decision trees, statistical learning},
  pdfauthor = {Torsten Hothorn, Achim Zeileis},
  colorlinks = {true},
  linkcolor = {Blue},
  citecolor = {Blue},
  urlcolor = {Red},
  hyperindex = {true},
  linktocpage = {true},
}

\begin{document}

\maketitle

\begin{abstract}
  The \pkg{partykit} package provides a flexible toolkit with infrastructure for
  learning, representing, summarizing, and visualizing a wide range of tree-structured
  regression and classification models. The functionality encompasses: (a)~Basic
  infrastructure for \emph{representing} trees (inferred by any algorithm) so that
  unified \code{print}/\code{plot}/\code{predict} methods are available.
  (b)~Dedicated methods for trees with \emph{constant fits} in the leaves (or terminal nodes)
  along with suitable coercion functions to create such tree models (e.g., by
  \pkg{rpart}, \pkg{RWeka}, PMML). (c)~A reimplementation of \emph{conditional inference
  trees} (\code{ctree}, originally provided in the \pkg{party} package). (d)~An extended
  reimplementation of \emph{model-based recursive partitioning} (\code{mob}, also originally
  in \pkg{party}) along with dedicated methods for trees with parametric models in the
  leaves. Here, a brief overview of the package and its design is given while more detailed
  discussions of items~(a)--(d) are available in vignettes accompanying the package.
\end{abstract}

\begin{keywords}
open-source software, recursive partitioning, regression trees, classification trees, decision trees, statistical learning. 
\end{keywords}

\section{Overview}

In the more than fifty years since \cite{Morgan+Sonquist:1963}
published their seminal paper on ``automatic interaction detection'',
a wide range of methods has been suggested that is usually
termed ``recursive partitioning'' or ``decision trees'' or
``tree(-structured) models'' etc. Particularly influential were the
algorithms CART \citep[classification and regression trees,][]{Breiman+Friedman+Olshen:1984},
C4.5 \citep{Quinlan:1993}, QUEST/GUIDE \citep{Loh+Shih:1997,Loh:2002},
and CTree \citep{Hothorn+Hornik+Zeileis:2006} among many others
\citep[see][for a recent overview]{Loh:2014}. Reflecting the heterogeneity
of conceptual algorithms, a wide range of computational implementations
in various software systems emerged: Typically the original authors of an
algorithms also provide accompanying software but many software systems,
e.g., including \pkg{Weka} \citep{Witten+Frank:2005} or \proglang{R} \citep{R},
also provide collections of various types of trees.
Within \proglang{R} the list of prominent packages includes
\pkg{rpart} \citep[implementing the CART algorithm]{rpart},
\pkg{RWeka} \citep[containing interfaces to J4.8, M5', LMT from \pkg{Weka}]{RWeka}, and
\pkg{party} \citep[implementing CTree and MOB]{party} among
many others. See the CRAN task view ``Machine Learning''
\citep{ctv} for an overview.

All of these algorithms and software implementations have to deal with
very similar challenges. However, due to the fragmentation of the
communities in which the corresponding research is published -- ranging
from statistics over machine learning to various applied fields -- many
discussions of the algorithms do not reuse established theoretical results
and terminology. Similarly, there is no common ``language'' for the
software implementations and different solutions are provided by
different packages (even within \proglang{R}) with relatively little
reuse of code.

The \pkg{partykit} tries to address the latter point and improve the
computational situation by providing a common unified infrastructure
for recursive partytioning in the \proglang{R} system for statistical
computing. In particular, \pkg{partykit}
provides tools for representing fitted trees along with printing,
plotting, and computing predictions. The design principles are:
\begin{itemize}
  \item One `agnostic' base class (\class{party}) which can encompass an extremely wide
        range of different types of trees.
  \item Subclasses for important types of trees, e.g., trees with constant
        fits (\class{constparty}) or with parametric models (\class{modelparty})
	in each terminal node (or leaf).
  \item Nodes are recursive objects, i.e., a node can contain child nodes.
  \item Keep (learning) data out of the recursive node and split structure.
  \item Basic printing, plotting, and predicting for raw node structure.
  \item Customization via suitable panel or panel-generating functions.
  \item Coercion from existing objects (\code{rpart}, \code{J48}, etc.) to the new class.
  \item Usage of simple/fast \proglang{S}3 classes and methods.
\end{itemize}
In addition to all of this generic infrastructure, two specific tree
algorithms are implemented in \pkg{partykit} as well: \fct{ctree}
for conditional inference trees \citep{Hothorn+Hornik+Zeileis:2006}
and \fct{mob} for model-based recursive partitioning \citep{Zeileis+Hothorn+Hornik:2008}.

\section{Installation and Documentation}

installation information

R-Forge

The vignette (\code{"partykit"}) introduces the basic \class{party} class
and associated infrastructure while three further vignettes discuss the tools
built on top of it: \code{"constparty"} covers the eponymous class for
constant-fit trees along with suitable coercion functions, and
\code{"ctree"} and \code{"mob"} discuss the new \fct{ctree} and
\fct{mob} implementations, respectively. Each of the vignettes can be viewed
within \proglang{R} via \code{vignette(}\emph{``name''}\code{, package = "partykit")}.

\section{User Interface}

formula interface

as.party

PMML

print, plot, predict

<<Titanic>>=
data("Titanic", package = "datasets")
ttnc <- as.data.frame(Titanic)
ttnc <- ttnc[rep(1:nrow(ttnc), ttnc$Freq), 1:4]
names(ttnc)[2] <- "Gender"
ttnc <- transform(ttnc, Treatment = factor(
  Gender == "Female" | Age == "Child", levels = c(FALSE, TRUE),
  labels = c("Normal\nMale&Adult", "Preferential\nFemale|Child")
))
@

<<rpart>>=
library("rpart")
ttnc_rp <- as.party(rpart(Survived ~ ., data = ttnc))
@

<<mob>>=
ttnc_mob <- glmtree(Survived ~ Treatment | Class + Gender + Age,
  data = ttnc, family = binomial, alpha = 0.05, prune = "BIC")
@

\setkeys{Gin}{width=\textwidth}
\begin{figure}[t!]
\centering
<<mob-plot, echo=FALSE, fig=TRUE, height=7, width=18>>=
grid.newpage()
pushViewport(viewport(layout = grid.layout(1, 2)))
pushViewport(viewport(layout.pos.col = 1, layout.pos.row = 1))
plot(ttnc_rp, newpage = FALSE, pop = TRUE,
  tp_args = list(reverse = FALSE, fill = grey(c(0.45, 0.8))))
popViewport()
pushViewport(viewport(layout.pos.col = 2, layout.pos.row = 1))
plot(ttnc_mob, newpage = FALSE, pop = TRUE, tp_args = list(ylines = 1.3,
  margins = c(1.5, 1, 1.5, 2.2), gp = gpar(fill = grey(c(0.45, 0.8))),
  linecol = "black"))
popViewport(2)
@
\caption{\label{fig} Tree visualizations of survival on Titanic:
\class{rpart} tree visualized by \pkg{partykit} (left) and
logistic-regression-based tree fitted by \fct{glmtree} (right).}
\end{figure}

\section{Internal Structure}

partysplit
partynode
party

constparty
modelparty

extractor functions

\section{Conclusions}

run-time, memory requirements, benchmarks?

\section*{Acknowledgments}

We are thankful to the organizers and participants of the ``Workshop on Classification and Regression Trees'' 2014,
sponsored by the Institute for Mathematical Sciences of the National University of Singapore,
for feedback and discussions.

\bibliography{../../vignettes/party.bib}

\end{document}