### 2014-03-25 pkg/partykit/inst/mloss/mloss2.Rnw
touch-ups in sections 1-3
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@553 255f4249-2263-4417-b1ea-8498779e27a7

\documentclass[a4paper]{article}
\usepackage{jmlr2e}
\usepackage{color,hyperref}

%% software markup
\let\proglang=\textsf
\let\pkg=\textit
\let\code=\texttt
\newcommand{\squote}[1]{`{#1}'}
\newcommand{\dquote}[1]{``{#1}''}
\newcommand{\fct}[1]{\texttt{#1}}
\newcommand{\class}[1]{\squote{\texttt{#1}}}

%% \usepackage{Sweave} is essentially
\RequirePackage[T1]{fontenc}
\RequirePackage{ae,fancyvrb}
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
\newenvironment{Schunk}{}{}
\setkeys{Gin}{width=.7\textwidth}
\SweaveOpts{engine=R, eps=FALSE, keep.source=TRUE, echo=FALSE, results=hide}
<<options, echo=FALSE>>=
options(prompt = "R> ", continue = "+  ", digits = 4, width = 70)
library("partykit")
set.seed(290875)
@

%% header declarations
\author{Torsten Hothorn \\ \addr{Universit\"at Z\"urich} \\ \addr{E-mail: Torsten.Hothorn@R-project.org} \AND
        Achim Zeileis \\ \addr{Universit\"at Innsbruck} \\ \addr{E-mail: Achim.Zeileis@R-project.org}}
\title{\pkg{partykit}: A Modular Toolkit for Recursive Partytioning in \proglang{R}}
\editor{(submission for MLOSS track, related to JMLR-07-240-1 that was handled by Cheng Soon Ong)}

% hyperref setup
\definecolor{Red}{rgb}{0.5,0,0}
\definecolor{Blue}{rgb}{0,0,0.5}
\hypersetup{%
  pdftitle = {partykit: An Open-Source Toolkit for Recursive Partytioning in R},
  pdfsubject = {submitted to JMLR-MLOSS},
  pdfkeywords = {open-source software, recursive partitioning, decision trees, statistical learning},
  pdfauthor = {Torsten Hothorn, Achim Zeileis},
  colorlinks = {true},
  linkcolor = {Blue},
  citecolor = {Blue},
  urlcolor = {Red},
  hyperindex = {true},
  linktocpage = {true},
}

\begin{document}

\maketitle

\begin{abstract}
  The \proglang{R} package \pkg{partykit} provides a flexible toolkit for
  learning, representing, summarizing, and visualizing a wide range of tree-structured
  regression and classification models. The functionality encompasses: (a)~Basic
  infrastructure for \emph{representing} trees (inferred by any algorithm) so that
  unified \code{print}/\code{plot}/\code{predict} methods are available.
  (b)~Dedicated methods for trees with \emph{constant fits} in the leaves (or terminal nodes)
  along with suitable coercion functions to create such trees (e.g., by
  \pkg{rpart}, \pkg{RWeka}, PMML). (c)~A reimplementation of \emph{conditional inference
  trees} (\code{ctree}, originally provided in the \pkg{party} package). (d)~An extended
  reimplementation of \emph{model-based recursive partitioning} (\code{mob}, also originally
  in \pkg{party}) along with dedicated methods for trees with parametric models in the
  leaves. Here, a brief overview of the package and its design is given while more detailed
  discussions of items~(a)--(d) are available in vignettes accompanying the package.
\end{abstract}

\begin{keywords}
open-source software, recursive partitioning, regression trees, classification trees, decision trees, statistical learning. 
\end{keywords}

\section{Overview}

In the more than fifty years since \cite{Morgan+Sonquist:1963}
published their seminal paper on ``automatic interaction detection'',
a wide range of methods has been suggested that is usually
termed ``recursive partitioning'' or ``decision trees'' or
``tree(-structured) models'' etc. Particularly influential were the
algorithms CART \citep[classification and regression trees,][]{Breiman+Friedman+Olshen:1984},
C4.5 \citep{Quinlan:1993}, QUEST/GUIDE \citep{Loh+Shih:1997,Loh:2002},
and CTree \citep{Hothorn+Hornik+Zeileis:2006} among many others
\citep[see][for a recent overview]{Loh:2014}. Reflecting the heterogeneity
of conceptual algorithms, a wide range of computational implementations
in various software systems emerged: Typically the original authors of an
algorithm also provide accompanying software but many software systems,
including \pkg{Weka} \citep{Witten+Frank:2005} or \proglang{R} \citep{R},
also provide collections of various types of trees.
Within \proglang{R} the list of prominent packages includes
\pkg{rpart} \citep[implementing CART]{rpart},
\pkg{RWeka} \citep[containing interfaces to J4.8, M5', LMT from \pkg{Weka}]{RWeka}, and
\pkg{party} \citep[implementing CTree and MOB]{party} among
many others. See the CRAN task view ``Machine Learning''
\citep{ctv} for an overview.

All of these algorithms and software implementations have to deal with
very similar challenges. However, due to the fragmentation of the
communities in which the corresponding research is published -- ranging
from statistics over machine learning to various applied fields -- many
discussions of the algorithms do not reuse established theoretical results
and terminology. Similarly, there is no common ``language'' for the
software implementations and different solutions are provided by
different packages (even within \proglang{R}) with relatively little
reuse of code.

The \pkg{partykit} aims at mitigating the latter issue
by providing a common unified infrastructure
for recursive partytioning in the \proglang{R} system for statistical
computing. In particular, \pkg{partykit}
provides tools for representing fitted trees along with printing,
plotting, and computing predictions. The design principles are:
\begin{itemize}
  \item One `agnostic' base class (\class{party}) which can encompass an extremely wide
        range of different types of trees.
  \item Subclasses for important types of trees, e.g., trees with constant
        fits (\class{constparty}) or with parametric models (\class{modelparty})
	in each terminal node (or leaf).
  \item Nodes are recursive objects, i.e., a node can contain child nodes.
  \item Keep the (learning) data out of the recursive node and split structure.
  \item Basic printing, plotting, and predicting for raw node structure.
  \item Customization via suitable panel or panel-generating functions.
  \item Coercion from existing object classes in \proglang{R} (\code{rpart}, \code{J48}, etc.) to the new class.
  \item Usage of simple/fast \proglang{S}3 classes and methods.
\end{itemize}
In addition to all of this generic infrastructure, two specific tree
algorithms are implemented in \pkg{partykit} as well: \fct{ctree}
for conditional inference trees \citep{Hothorn+Hornik+Zeileis:2006}
and \fct{mob} for model-based recursive partitioning \citep{Zeileis+Hothorn+Hornik:2008}.

\section{Installation and Documentation}

The \pkg{partykit} package is an add-on package for the \proglang{R} system
for statistical computing. It is available from the Comprehensive \proglang{R}
Archive Network (CRAN) at \url{http://CRAN.R-project.org/package=partykit} and
can be easily installed in the usual ways from within \proglang{R}, e.g., using
\fct{install.packages}. It depends on \proglang{R} (at least 2.15.0) as well
as the base packages \pkg{graphics}, \pkg{grid}, \pkg{stats}, and the recommended
\pkg{survival}. Furthermore, various suggested packages are needed
for certain special functionalities in the package. To install all of these
required and suggested packages in one go, the command
\code{install.packages("partykit", dependencies = TRUE)} can be used.

In addition to the stable release version on CRAN, the
current development release can be installed from \proglang{R}-Forge
\citep{Theussl+Zeileis:2009}. In addition to source and binary packages
the entire version history is available through
\proglang{R}-Forge's \pkg{Subversion} source code management system.

Along with the package extensive documentation with examples is shipped.
The manual pages provide basic technical information on all functions while
much more detailed descriptions along with hands-on examples are provided in
the four package vignettes. First, the vignette \code{"partykit"}
introduces the basic \class{party} class
and associated infrastructure while three further vignettes discuss the tools
built on top of it: \code{"constparty"} covers the eponymous class (as well as
the simplified \class{simpleparty} class) for constant-fit trees along with
suitable coercion functions, and \code{"ctree"} and \code{"mob"} discuss the
new \fct{ctree} and \fct{mob} implementations, respectively. Each of the
vignettes can be viewed within \proglang{R} via
\code{vignette(}\emph{``name''}\code{, package = "partykit")} and the underlying
source code (in \proglang{R} with {\LaTeX} text) is also available in the source
package.

\section{User Interface}

The \pkg{partykit} package provides functionality at different levels. First,
there is basic infrastructure for representing, modifying, and displaying trees
and recursive partitions -- these tools are mostly intended for developers
and described in the next section. Second, there are tools for inferring trees
from data or for importing trees inferred by other software into \pkg{partykit}.

\begin{table}
\centering
\begin{tabular}{llll}
\hline
Algorithm  & Software                                                & Object class       & Original reference \\ \hline
CART/RPart & \code{rpart::rpart} + \code{as.party}                   & \code{constparty}  & \cite{Breiman+Friedman+Olshen:1984} \\
C4.5/J4.8  & \pkg{Weka}/\code{RWeka::J48} + \code{as.party}          & \code{constparty}  & \cite{Quinlan:1993} \\
QUEST      & \proglang{SPSS}/\pkg{AnswerTree} + \code{pmmlTreeModel} & \code{simpleparty} & \cite{Loh+Shih:1997} \\
CTree      & \code{ctree}                                            & \code{constparty}  & \cite{Hothorn+Hornik+Zeileis:2006} \\
MOB        & \code{mob}, \code{lmtree}, \code{glmtree}, \dots        & \code{modelparty}  & \cite{Zeileis+Hothorn+Hornik:2008} \\
EvTree     & \code{evtree::evtree}                                   & \code{constparty}  & \cite{evtree} \\ \hline
\end{tabular}
\caption{\label{tab:overview} Selected tree algorithms than can be interfaced through \pkg{partykit}. The
  software column lists external software, \proglang{R} functions from other packages (with \code{::} syntax)
  and from \pkg{partykit}.}
\end{table}

While originally an important goal for the development of \pkg{partykit} was to
provide infrastructure for the authors' own tree induction algorithms
CTree and MOB,
the design was very careful to separate as much functionality as possible into
more general classes that are useful for a far broader class of tree algorithms.
In particular, to be able to print/plot/predict different trees in a unified way,
there are so-called coercion functions for transforming trees learned in other
software packages (inside and outside of \proglang{R}) into the classes provided
by \pkg{partykit}. Specifically, tree objects learned by \fct{rpart} from the
package of the same name (\citealp{rpart}, implementing CART, \citealp{Breiman+Friedman+Olshen:1984})
and by \fct{J48} from \pkg{RWeka} (\citealp{RWeka}, interfacing \pkg{Weka}'s J4.8
algorithm for C4.5, \citealp{Quinlan:1993}) can be coerced by \fct{as.party}
to the same object class \class{constparty}. This is a general class that can
in principle represent all the major classical tree algorithms with constant
fits in the terminal nodes. Also, the same class is employed for conditional
inference trees (CTree) that can be learned with the \fct{ctree} function
directly within \pkg{partykit} or evolutionary trees from package \pkg{evtree}
\citep{evtree}.

Not only trees learned within \proglang{R} can be transformed to the
proposed infrastructure but also trees from other software packages. Either
a dedicated interface has to be created using the building blocks described
in the next section (e.g., as done for the J4.8 tree in \pkg{RWeka}) or
PMML (Predictive Model Markup Language) can be used as an intermediate
exchange format. This is an XML standard created by an international
consortium \citep{DMG:2014} that includes a \code{<TreeModel>} tag with
support for constant-fit classification and regression trees. The function
\fct{pmmlTreeModel} allows to read these files and represents them as
\class{simpleparty} objects in \pkg{partykit}. The reason for not using the
\class{constparty} class as above is that the PMML format only stores
point predictions (e.g., a mean or proportion) rather than all observations
from the learning sample. So far, the PMML interface has been tested with
output from the \proglang{R} package \pkg{pmml} and \proglang{SPSS}'s
\pkg{AnswerTree} model. The latter includes an implementation of the
QUEST algorithm \citep{Loh+Shih:1997}.

Finally, the \pkg{partykit} function \fct{mob} implements model-based
recursive partitioning (MOB) along with ``mobster'' interfaces for certain
models (e.g., \fct{lmtree}, \fct{glmtree}). These return objects of class
\class{modelparty} where nodes are associated with statistical models (as
opposed to simple constant fits). In principle, this may also be adapted
to other model trees (such as GUIDE, LMT, or M5') but no such interface
is currently available.

All of these different trees (see Table~\ref{tab:overview} for an overview)
use the same infrastructure at the core but possibly with different options
enabled. In all cases, the functions \fct{print}, \fct{plot}, and \fct{predict}
can be used to create textual and graphical displays of the tree and for
computing predictions on new data, respectively. As an example for the
visualizations, Figure~\ref{fig:ttnc} shows two different trees fitted
to the well-known data on survival of passengers on the ill-fated maiden
voyage of the RMS~Titanic: The left panel shows a CART tree with constant fits
learned by \pkg{rpart} and converted to \pkg{partykit}. The right panel
shows a MOB tree learned within \pkg{partykit} with a logistic regression
model for treatment effects in the terminal nodes. Additionally, the
are further utility functions, e.g.,
\fct{nodeapply} can be employed to access further information
stored in the nodes of a tree and \fct{nodeprune} can prune selected nodes.

<<Titanic>>=
data("Titanic", package = "datasets")
ttnc <- as.data.frame(Titanic)
ttnc <- ttnc[rep(1:nrow(ttnc), ttnc$Freq), 1:4]
names(ttnc)[2] <- "Gender"
ttnc <- transform(ttnc, Treatment = factor(
  Gender == "Female" | Age == "Child", levels = c(FALSE, TRUE),
  labels = c("Normal\nMale&Adult", "Preferential\nFemale|Child")
))
@

<<rpart>>=
library("rpart")
ttnc_rp <- as.party(rpart(Survived ~ ., data = ttnc))
@

<<mob>>=
ttnc_mob <- glmtree(Survived ~ Treatment | Class + Gender + Age,
  data = ttnc, family = binomial, alpha = 0.05, prune = "BIC")
@

\setkeys{Gin}{width=\textwidth}
\begin{figure}[t!]
\centering
<<mob-plot, echo=FALSE, fig=TRUE, height=7, width=18>>=
grid.newpage()
pushViewport(viewport(layout = grid.layout(1, 2)))
pushViewport(viewport(layout.pos.col = 1, layout.pos.row = 1))
plot(ttnc_rp, newpage = FALSE, pop = TRUE,
  tp_args = list(reverse = FALSE, fill = grey(c(0.45, 0.85))))
popViewport()
pushViewport(viewport(layout.pos.col = 2, layout.pos.row = 1))
plot(ttnc_mob, newpage = FALSE, pop = TRUE, tp_args = list(ylines = 1.3,
  margins = c(1.5, 1, 1.5, 2.2), gp = gpar(fill = grey(c(0.45, 0.85))),
  linecol = "black"))
popViewport(2)
@
\caption{\label{fig:ttnc} Tree visualizations of survival on Titanic:
\class{rpart} tree converted with \fct{as.party} and visualized by \pkg{partykit} (left); and
logistic-regression-based tree fitted by \fct{glmtree} (right).}
\end{figure}

\section{Developer Infrastructure}

especially for developers who either want to implement new tree algorithms
or represent trees learning in other systems in \proglang{R} to facilitate
predictions, comparisons, etc.

partysplit
partynode
party

constparty
simpleparty
modelparty

extractor functions

\section{Conclusions}

run-time, memory requirements, benchmarks?

\section*{Acknowledgments}

We are thankful to the organizers and participants of the ``Workshop on Classification and Regression Trees'' 2014,
sponsored by the Institute for Mathematical Sciences of the National University of Singapore,
for helpful feedback and stimulating discussions.

\newpage

\bibliography{../../vignettes/party.bib}

\end{document}