### 2014-10-22 pkg/partykit/inst/mloss/letter_mloss3.tex
prepare p2p and letter
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@627 255f4249-2263-4417-b1ea-8498779e27a7

\documentclass[ngerman,uzh]{scrlttr2}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

%% ANPASSUNGSMÖGLICHKEITEN
\KOMAoptions{%
%backaddress=false%  Keine Rücksendeadresse im Adressfeld
%,foldmarks=false%  Faltmarken ausschalten
}

\usepackage{babel}

\begin{document}

%% VARIABLEN
% Falls Unterschrift vom definierten Namen abweicht:
\setkomavar{signature}{Torsten Hothorn}
\setkomavar{subject}{Betreff}
%\setkomavar{specialmail}{Interne Post}

\begin{letter}{
Dr.~Cheng Soon Ong \\
MLOSS Editor \\
Journal of Machine Learning Research}

\subject{JMLR-14-133-3 revision}

\opening{Dear Cheng Soon,}

I uploaded a rebuttal letter for our manuscript
``\textbf{partykit}: A Modular Toolkit for Recursive Partytioning in R'' to the
Machine Learning Open-Source Software track of the Journal of Machine
Learning Research.

Referee 3 criticised the limited breath of our contribution and suggested to
copy the \texttt{cforest} function from package \emph{party} to package
\emph{partykit}, iterating an argument by Referee 2 from the previous
round. In the rebuttal letter, we explain the reason why it is currently not
helpful to move this function from one CRAN package to another one. In
addition, we took the liberty to discuss why we think that \emph{partykit}
is actually much more general than \emph{party} and, if one adopts a certain
point of view, also more general than most of the established machine
learning toolboxes. Since no changes in the manuscript were requested, we
did not touch the four page description.

Sincerely yours, \\
\includegraphics[width = .25\textwidth]{TH_Unterschrift}

\end{letter}

\end{document}