### 2014-11-26 www_src_partykit/03_getting_started.Rmd
add citations
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@638 255f4249-2263-4417-b1ea-8498779e27a7

---
layout: page
title: Getting Started
permalink: /getting_started/
---

Your first regression tree
========================================================

First, install and fire-up R on your computer. Within R, one needs to
install the party package by typing

```{r eval=FALSE}
install.packages("partykit")
```

and hitting the ENTER key. Once the package is installed, you can
load it using

```{r results: hide}
library("partykit")
```

Now all party functions are ready to be used, for example the ctree()
function for fitting a regression tree to the Ozone data (after removing observations
with missing response):

```{r}
### regression
airq <- subset(airquality, !is.na(Ozone))
airct <- ctree(Ozone ~ ., data = airq,   
               control = ctree_control(maxsurrogate = 3))
airct
```

The tree is represented by an object called airct which can be plotted

```{r fig.width=9, fig.height=6}
plot(airct)
```

or used for computing predictions

```{r}
summary(predict(airct))
```

which can be compared to the actual response values:

```{r fig.width=6, fig.height=6}
plot(airq$Ozone, predict(airct))
abline(a = 0, b = 1)
```