### 2014-11-26 www_src_partykit/04_publications.Rmd

---
layout: page
title: Publications
permalink: /publications/
---

Scholary papers describing the methodology
========================================================

```{r setup, include = FALSE, cache = FALSE}
library("RefManageR")
bib <- ReadBib("TH.bib", check = FALSE)
BibOptions(check.entries = FALSE, style = "html", cite.style = "authoryear",
           bib.style = "numeric")
```
The partykit package implements conditional inference trees
`r AutoCite(bib, "Hothorn:2006:JCGS", before = "function ctree(), ")`,
and model-based recursive partitioning 
`r AutoCite(bib, "Zeileis+Hothorn+Hornik:2008", before = "function mob(), ")`.

**References**

```{r results = "asis", echo = FALSE}
PrintBibliography(bib, .opts = list(check.entries = FALSE, sorting = "ynt"))
```