### 2014-11-28 pkg/partykit/inst/mloss/review4/p2p.tex
4th round
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@643 255f4249-2263-4417-b1ea-8498779e27a7

The paper itself is fine.  Last review I had no further comments on the paper itself.  My review here then is restricted to the point-by-point reply.  This "discussion" is going downhill quickly and I will thus keep my comments short.

The authors have pointed out some measures of project activity, which are fine.  Of my concerns, the activity level and lack of open mailing list are less of a big deal, and my review stated as much.

The remaining sticking point is this:

"I am not convinced that the breadth of this contribution is sufficient." (from the previous review)

The author's replies have not convinced me otherwise despite their extensive length.  My suggestion was that support for forests be added, or some other support be added to increase the utility or breadth of partykit.  The responses given to this idea by the authors are concluded with remarks like the following:

"These [ideas] deserve further thought and work, both conceptually and computationally."

"Therefore, we would much rather follow Reviewer 2 and devote more future work to this important topic."

"This is on our to-do list, though."

If these things are on the authors' to-do lists, then they should do them and resubmit when they are done.

Personally I am not willing to consider acceptance as-is with the assumption that these features will be added later because the authors have stated elsewhere that partykit is mature and now only requires maintenance.

The package would be stronger and relevant to a wider audience, even if the added support was only preliminary.  And according to the authors themselves, this would be a significant contribution because flexible support for forests is a "real challenge".