### 2015-01-06 pkg/partykit/inst/mloss/letter_mloss4.tex
add letter
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@678 255f4249-2263-4417-b1ea-8498779e27a7

\documentclass[ngerman,uzh]{scrlttr2}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

%% ANPASSUNGSMÖGLICHKEITEN
\KOMAoptions{%
%backaddress=false%  Keine Rücksendeadresse im Adressfeld
%,foldmarks=false%  Faltmarken ausschalten
}

\usepackage{babel}

\begin{document}

%% VARIABLEN
% Falls Unterschrift vom definierten Namen abweicht:
\setkomavar{signature}{Torsten Hothorn}
\setkomavar{subject}{Betreff}
%\setkomavar{specialmail}{Interne Post}

\begin{letter}{
Dr.~Cheng Soon Ong \\
MLOSS Editor \\
Journal of Machine Learning Research}

\subject{JMLR-14-133-4 revision}

\opening{Dear Cheng Soon,}

I uploaded a revised version of the \textbf{partykit} package (version
0.8-4, also the current CRAN version). Following the request by referee 3,
we implemented \textbf{cforest()} in this version. The benefits and
shortcomings of this implementation of random forests, compared to the
implementations in the \textbf{randomForest} and \textbf{party} packages, are
discussed in the rebuttal.

The manuscript was not changed and thus I did not upload a new version.

Best regards, \\
\includegraphics[width = .25\textwidth]{TH_Unterschrift}

\end{letter}

\end{document}